---
title: "Class Reference"
author: "Molecule Jongwilai"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

### How to set the working directory

The command to set a working directory is setwd(). For example, this sets the working directory to a directory called data_journalism_2024_fall inside a directory called jour472 located inside your home directory (the ~ is equivalent of /Users/[youruser] on a Mac).

```{r}
setwd("~/jour472/data_journalism_2024_fall")
```

### How to install and load the tidyverse

```{r}
install.packages("tidyverse")
library(tidyverse)
```

Here is how to import the data set (.rds file).

```{r}
maryland_winred_contributions <- read_rds("maryland_winred.rds")
```

Here is how to view the number of columns and roles, a list of the columns, the data type of each column and first few values of each column. 

```{r}
glimpse(maryland_winred_contributions)
```
Here is how to print out columns and first six rows of data.

```{r}
head(maryland_winred_contributions)
```
Here is how to group the same individual values of the column and count their number.

```{r}
maryland_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    count_contributes = n()
  )

```
From the previous function, here is how to arrange the list of value in descending order. 

```{r}
maryland_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    count_contributes = n()
  ) |>
  arrange(desc(count_contributes))
```
From the previous function, here is how to group the values of more than one column. 

```{r}
maryland_winred_contributions |>
  group_by(memo_text, date) |>
  summarise(
    count_contributes = n()
  ) |>
  arrange(desc(count_contributes))
```
From the previous function, here is how to group the values of the column, count their number, and add more function to sum up another countable value from different column of each group of initial value. 

```{r}
maryland_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    count_contributes = n(),
    total_amount = sum(amount)
  ) |>
  arrange(desc(count_contributes))
```
From the previous function, here is how to group the values of the column, count their number, and add more functions to find total amount, mean, media of another countable value from different column of each group of initial value.

```{r}
maryland_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    count_contributes = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    media_amount = median(amount)
  ) |>
  arrange(desc(count_contributes))
```
From the previous function, here is how to group the values of the column, count their number, and add more functions to find total amount, mean, media, min, max of another countable value from different column of each group of initial value.

```{r}
maryland_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    count_contributes = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    media_amount = median(amount),
    min_amount = min(amount),
    max_amount = max(amount)
  ) |>
  arrange(desc(count_contributes))
```
Here is how to sort rows in the data set from a column's values in max to min amount order. 

```{r}
maryland_winred_contributions |>
  arrange(desc(amount))
```