---
title: "lab_03"
author: "Molecule Jongwilai"
date: "2024-09-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse
library(tidyverse)

# Load janitor
library(janitor)
```

## Load data

```{r}
# Load MD slice of FEC House candidate contribution data
md_contribs <- read_rds("data/lab_03.rds")
```

## Fix header issues

You'll need to fix one issue with the column names using janitor. Write a code block to do that, making sure that you save the results to the same dataframe.

```{r}
md_contribs <- md_contribs |>
  clean_names()
```


## Answer questions

Q1: Which seven cities in Maryland had the largest *number* of contributions from donors? How many did each have? To answer this question, you'll have to clean the messy "contributor_city" field.

Note: To answer this question, you'll need to see every variation of the "contributor_city" for the top seven cities, which you should identify through some initial analysis. Then find variations of them and change those variations into the most common version.

A1: The seven cities that have the largest number of contributions from donors are;
1. Bethesda with 1,177 contributions
2. Chevy Chase with 704 contributions
3. Baltimore with 691 contributions
4. Annapolis with 492 contributions 
5. Potomac with 491 contributions
6. Silver Spring with 408 contributions
7. Rockville with 359 contributions 

```{r}
md_contribs |>
  get_dupes()
```
```{r}
cities <- md_contribs |>
  group_by(contributor_city) |>
  summarise(
    count_city = n()
  )
```
```{r}
md_contribs <- md_contribs |>
  mutate(contributor_city = str_to_title(contributor_city)) |>
  mutate(contributor_city = case_when(
    contributor_city == "Anapolis" ~ "Annapolis",
    contributor_city == "Balrimore" ~ "Baltimore",
    contributor_city == "Potamic" ~ "Potomac",
    contributor_city == "Rockwell" ~ "Rockville",
    contributor_city == "Silver Springs" ~ "Silver Spring",
    TRUE ~ contributor_city 
  ))
```
```{r}
clean_cities <- md_contribs |>
  group_by(contributor_city) |>
  summarise(
    count_city = n()
  ) |>
  arrange(desc(count_city))
```


Q2: Find the 5-digit ZIP code in Maryland that had the highest total amount of contributions (by dollar value). You will need to use mutate to ensure that you have a zip code column with the first five characters from the existing zip code column.

Where is that zip code and what kind of area is it, demographically? Which committee is the largest recipient of money from that zip code? How many of the top five recipients are Maryland candidates?

A2:
- 20817 is the 5-digit ZIP code that had the highest total amount of contributions with $731,489.80

- 20817 is the ZIP code for Bethesda area in Montgomery County, located in central Maryland. According to census.gov (https://www.census.gov/quickfacts/fact/table/bethesdacdpmaryland,US/PST045223), Bethesda has population of 68,056 from 2021 census, 73.4% of white alone population, $185,546 median household income in 2018 - 2022, $116,088 income per capita in 2018 - 2022, and 3.9% of persons in poverty.

- HELMER FOR CONGRESS is the committee that receives the largest amount of money with $17,878.00 in the 20817 area.

- Among the top five recipients, there are 2 Maryland candidates which are Jamie Raskin, U.S. representative for Maryland's 8th congressional district, and April McClain Delaney, the Democratic party's nominee for Maryland's 6th congressional district in the 2024 election.

```{r}
md_contribs <- md_contribs |>
  mutate(contributor_zip = str_sub(contributor_zip, start=1L, end=5L))
```
```{r}
md_contribs |>
  group_by(contributor_zip) |>
  summarise(
    total_amount_of_contributions = sum(contribution_receipt_amount)
  ) |>
  arrange(desc(total_amount_of_contributions))
```
```{r}
md_contribs |>
  filter(contributor_zip == "20817") |>
  group_by(committee_name) |>
  summarise(
    each_committee_amount = sum(contribution_receipt_amount)
  ) |>
  arrange(desc(each_committee_amount))
```
Q3: Let's look at top individual donors. Write code to generate a list of top donors using name and city, adding up the total amount of contributions. What does the original data say about the top two donors - who are they and what do they do? Do you see anything in the list of top 10 donors that would make you uncertain about the accuracy of these results?

A3:
- The top two donors are;
1. Melanie Jacobsen, a farmer from Bethesda (*who donated on January 12, 2024, 26 times). 
2. Mark P. Jacobsen, a Banker, and CEO from Bethesda (*who donated on January 12, 2024, 18 times).

- There are 2 issues I spot that make me think this result may not be completely accurate;
1. The 2nd top donor, Mark P. Jacobsen, and 5th, Mark Jacobsen, may be the same person as they are from the same city and have the same occupation. 
2. The 3rd top donor, Jeffrey L. Drezner, and 7th, Jeffrey Drezner, may be the same person as they are from the same city and have the same occupation. 
Hence, I might do a bit more research to confirm my suspicion before correcting the data.  

```{r}
top_donors <- md_contribs |>
  group_by(contributor_name, contributor_city) |>
  summarise(
    amount_of_contributions = sum(contribution_receipt_amount)
  ) |>
  arrange(desc(amount_of_contributions))
```