---
title: "Data Analysis Project"
author: "Group 4"
date: "11-16-2024"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Deliverable 2 (week 12) - 10 percent*: 

*Question 1* Lists your questions - if i asked you to refine them, show me that you've done that.

*Answer 1* 
Q1. For the big picture, in *Maryland*,
1.1 How many voters registered for each party in the 2024 primary and general election? 
1.1.1 How have these numbers changed compared to those in previous elections?
1.2 Compared to the total population of the state, what is the proportion of voters who register for the election?
1.3 Compared to the 2024 presidential election result, does the registration pattern match with the result? 
---
Q2. At the *Congressional District level*
2.1 How many voters registered for each party in the 2024 primary and general election?
2.1.1 How have these numbers changed compared to those in previous elections?
2.2 Are there any patterns for voter registration from every district?
2.2.1 Are there any districts that don't fall into this pattern? 
2.2.2 How are those districts different from others in terms of location, demography, economy, etc.?
2.3 Compared to the total population of each district, what is the proportion of voters who register for the election? 
2.3.1 What are the differences between precincts having large and small proportions of voter registration?
2.4 Compared to the representative race result, does the registration pattern match with the result? 
---
Q3. At the *Legislative Districts level*
3.1 How many voters registered for each party in the 2024 primary and general election?
3.1.1 How have these numbers changed compared to those in previous elections?
3.2 Are there any patterns for voter registration from every district?
3.2.1 Are there any districts that don't fall into this pattern? 
3.2.2 How are those districts different from others in terms of location, demography, economy, etc.?
3.3 Compared to the total population of each district, what is the proportion of voters who register for the election? 
3.3.1 What are the differences between precincts having large and small proportions of voter registration?
3.4 Compared to the election result, does the registration pattern match with the result? 
---
Q4. At the *Precinct level*
4.1 How many voters registered for each party in the 2024 primary and general election?
4.1.1 How have these numbers changed compared to those in previous elections?
4.2 Are there any patterns for voter registration from every precinct?
4.2.1 Are there any precincts that don't fall into this pattern? 
4.2.2 How are those precincts different from others in terms of location, demography, economy, etc.?
4.3 Compared to the total population of each precinct, what is the proportion of voters who register for the election? 
4.3.1 What are the differences between precincts having large and small proportions of voter registration? 
4.4 Compared to the presidential election result, does the registration pattern match with the result?
---
Q5. At the *County level*
5.1 How many voters registered for each party in the 2024 primary and general election?
5.1.1 How have these numbers changed compared to those in previous elections?
5.2 Are there any patterns for voter registration from every county?
5.2.1 Are there any counties that don't fall into this pattern? 
5.2.2 How are those counties different from others in terms of location, demography, economy, etc.?
5.3 Compared to the total population of each county, what is the proportion of voters who register for the election? 
5.3.1 What are the differences between counties having large and small proportions of voter registration? 
5.4 Compared to the 2024 presidential election result, does the registration pattern match with the result? 

*Question 2* Loads and cleans the core data set to prepare for analysis.

*Answer 2*

```{r}
#Load Libraries 
library(tidyverse)
library(janitor)
library(tidycensus)
library(rvest)
library(dplyr)
```


```{r}
#Prep dataframe for Q1 focusing on Maryland

#Registration Data
md_primary_24 <- read_csv("data/county/county_primary_24.csv")
md_general_24 <- read_csv("data/county/county_general_24.csv")
md_primary_20 <- read_csv("data/county/county_primary_20.csv")
md_primary_20 <- md_primary_20 |>
  rename(County = 1)
md_general_20 <- read_csv("data/county/county_general_20.csv")

#Population Data
state_population <- get_acs(geography="state", 
                       variables = "B01003_001", 
                       year=2022)
md_population <- state_population |>
  slice(21)

#Presidential Election Result
pres_election_result_url <- "https://elections.maryland.gov/elections/2024/general_Results/gen_results_2024_1.html"
pres_election_result_html <- pres_election_result_url |>
  read_html() |>
  html_table()
pres_election_result <- pres_election_result_html[[1]]

```
```{r}
#Prep dataframe for Q2 focusing on congressional district level

#Registration Data
congress_primary_24 <- read_csv("data/congress_level/congress_primary_24.csv")
congress_general_24 <- read_csv("data/congress_level/congress_general_24.csv")
congress_primary_20 <- read_csv("data/congress_level/congress_primary_20.csv")
congress_primary_20 <- congress_primary_20 |>
  rename(District = 1) |>
  rename(county = 2) |>
  rename(BAR = 3) |>
  rename(DEM = 4) |>
  rename(OTH = 5) |>
  rename(REP = 6) |>
  rename(UNA = 7) |>
  rename(WCP = 8) |>
  rename(total = 9) 
congress_general_20 <- read_csv("data/congress_level/congress_general_20.csv")

#Population Data
congress_population <- get_acs(geography="congressional district", 
                       variables = "B01003_001", state = "MD", 
                       year=2022)

#Representative Election Result
representative_election_result_url <- "https://elections.maryland.gov/elections/2024/general_Results/gen_results_2024_3.html"
representative_election_result_html <- representative_election_result_url |>
  read_html() |>
  html_table()
representative_election_result1 <- representative_election_result_html[[1]]
representative_election_result2 <- representative_election_result_html[[2]]
representative_election_result3 <- representative_election_result_html[[3]]
representative_election_result4 <- representative_election_result_html[[4]]
representative_election_result5 <- representative_election_result_html[[5]]
representative_election_result6 <- representative_election_result_html[[6]]
representative_election_result7 <- representative_election_result_html[[7]]
representative_election_result8 <- representative_election_result_html[[8]]
```
```{r}
#Prep dataframe for Q3 focusing on legislative district level

#Registration Data
legis_primary_24 <- read_csv("data/legis_level/legis_primary_24.csv")
legis_general_24 <- read_csv("data/legis_level/legis_general_24.csv")
legis_primary_20 <- read_csv("data/legis_level/legis_primary_20.csv")
legis_general_20 <- read_csv("data/legis_level/legis_general_20.csv")

#Population Data
legis_population <- get_acs(geography="state legislative district (lower chamber)", 
                       variables = "B01003_001", state = "MD", 
                       year=2022)

#Presidential Election Result 
#Download Precinct Result and Code
precint_result <- read_csv("data/legis_level/result/result.csv")
legis_code <- read_csv("data/legis_level/result/code.csv")
#Prep result dataframe
precint_result <- precint_result |>
  filter(Contest == "President and Vice President of the United States")
precint_result <- precint_result |>
  select(2,4,6,7) |>
  clean_names() |>
  mutate(precinct_name = as.numeric(gsub("-","", precinct_name)))
#Prep code dataframe
legis_code <- legis_code |>
  select(3,4,10) |>
  filter(!is.na(ED_PRECINCT)) |>
  filter(!is.na(LEGISLATIVE_SUB_DISTRICT_CODE))
#Join result dataframe with code dataframe to generate legislative district results 
legis_result_draft <- precint_result |> 
  left_join(legis_code, join_by(jurisdiction==COUNTY_NAME, precinct_name==ED_PRECINCT)) |>
  filter(!is.na(LEGISLATIVE_SUB_DISTRICT_CODE)) |>
  group_by(LEGISLATIVE_SUB_DISTRICT_CODE, party) |>
  summarise(
    total_votes = sum(total_votes))
#Reshape dataframe
legis_result_DEM <- legis_result_draft |>
  filter(party == "DEM") |>
  select(1,3) |>
  rename(DEM_result = 2)
legis_result_GRN <- legis_result_draft |>
  filter(party == "GRN") |>
  select(3) |>
  rename(GRN_result = 2)
legis_result_LIB <- legis_result_draft |>
  filter(party == "LIB") |>
  select(3) |>
  rename(LIB_result = 2)
legis_result_NON <- legis_result_draft |>
  filter(party == "NON") |>
  select(3) |>
  rename(NON_result = 2)
legis_result_REP <- legis_result_draft |>
  filter(party == "REP") |>
  select(3) |>
  rename(REP_result = 2)
legis_result_UNF <- legis_result_draft |>
  filter(party == "UNF") |>
  select(3) |>
  rename(UNF_result = 2)
legis_result <- bind_cols(legis_result_DEM, legis_result_GRN, legis_result_LIB, legis_result_NON, legis_result_REP, legis_result_UNF) |>
  select(-3,-5,-7,-9,-11) |>
  rename(legislative_district_code = 1)

```

```{r}
#Prep dataframe for Q4 focusing on legislative precinct level

#Presidential Result Data
precinct_result <- read_csv("data/precinct/precinct result.csv")
precinct_result_DEM <- precinct_result |>
  select(1, 2, 3, 4, 6, 7) |>
  filter(Contest == "President and Vice President of the United States") |>
  select(-1) |>
  filter(Party == "DEM") |>
  rename(total_votes_DEM = 5) |>
  select(-4)
precinct_result_REP <- precinct_result |>
  select(1, 2, 3, 4, 6, 7) |>
  filter(Contest == "President and Vice President of the United States") |>
  select(-1) |>
  filter(Party == "REP") |>
  rename(total_votes_REP = 5) |>
  select(-1,-2,-3,-4)
precinct_result_LIB <- precinct_result |>
  select(1, 2, 3, 4, 6, 7) |>
  filter(Contest == "President and Vice President of the United States") |>
  select(-1) |>
  filter(Party == "LIB") |>
  rename(total_votes_LIB = 5) |>
  select(-1,-2,-3,-4)
precinct_result_GRN <- precinct_result |>
  select(1, 2, 3, 4, 6, 7) |>
  filter(Contest == "President and Vice President of the United States") |>
  select(-1) |>
  filter(Party == "GRN") |>
  rename(total_votes_GRN = 5) |>
  select(-1,-2,-3,-4)
precinct_result_UNF <- precinct_result |>
  select(1, 2, 3, 4, 6, 7) |>
  filter(Contest == "President and Vice President of the United States") |>
  select(-1) |>
  filter(Party == "UNF") |>
  rename(total_votes_UNF = 5) |>
  select(-1,-2,-3,-4)
precinct_result_NON <- precinct_result |>
  select(1, 2, 3, 4, 6, 7) |>
  filter(Contest == "President and Vice President of the United States") |>
  select(-1) |>
  filter(Party == "NON") |>
  rename(total_votes_NON = 5) |>
  select(-1,-2,-3,-4)
precinct_result_cleaned <- bind_cols(precinct_result_DEM, precinct_result_REP, precinct_result_LIB, precinct_result_GRN, precinct_result_UNF, precinct_result_NON)
precinct_result_final <- precinct_result_cleaned |>
  mutate(total_votes = total_votes_DEM + total_votes_REP + total_votes_LIB + total_votes_GRN + total_votes_UNF + total_votes_NON)
```


```{r}
#Prep dataframe for Q5 focusing on county level

#Registration Data
county_primary_24 <- read_csv("data/county/county_primary_24.csv")
county_general_24 <- read_csv("data/county/county_general_24.csv")
county_primary_20 <- read_csv("data/county/county_primary_20.csv")
county_primary_20 <- county_primary_20 |>
  rename(County = 1)
county_general_20 <- read_csv("data/county/county_general_20.csv")

#Population Data
county_population <- get_acs(geography="county", 
                       variables = "B01003_001", 
                       state = "MD", year=2019)

#Presidential Election Result (county level)
pres_election_result_county_url <- "https://elections.maryland.gov/elections/2024/general_Results/gen_detail_results_2024_1_1.html"
pres_election_result_county_html <- pres_election_result_county_url |>
  read_html() |>
  html_table()
pres_election_result_county1 <- pres_election_result_county_html[[1]]
pres_election_result_county2 <- pres_election_result_county_html[[2]]
pres_election_result_county2 <- pres_election_result_county2|>
  select(-1)
pres_election_result_county3 <- pres_election_result_county_html[[3]]
pres_election_result_county3 <- pres_election_result_county3|>
  select(-1)
pres_election_result_county4 <- pres_election_result_county_html[[4]]
pres_election_result_county4 <- pres_election_result_county4|>
  select(-1)
pres_election_result_county5 <- pres_election_result_county_html[[5]]
pres_election_result_county5 <- pres_election_result_county5|>
  select(-1)
pres_election_result_county6 <- pres_election_result_county_html[[6]]
pres_election_result_county6 <- pres_election_result_county6|>
  select(-1)
pres_election_result_county <- bind_cols(pres_election_result_county1, pres_election_result_county2, pres_election_result_county3, pres_election_result_county4, pres_election_result_county5, pres_election_result_county6)
```
*Question 3* Shows basic exploratory analysis to demonstrate an understanding of the dataset, including the number of rows and columns, any obvious limitations or flaws and any reasons why it might not be able to answer the questions you've posed last week.

*Answer 3*
Q1
We may not be able to compare the change in the number of registrations for LIB and NLM parties (from primary), and NLM parties (from general) between 2024 and 2020 as the dataframes from 2020 don’t have data for these parties.
Q2
We may not be able to compare the change in the number of registrations at the congressional level for LIB and NLM parties (from primary), and NLM party (from general) between 2024 and 2020 as the dataframes from 2020 don’t have data for these parties. 
Q3
We may not be able to compare the change in the number of registrations at the legislative level for the NLM party (from both primary and general) between 2024 and 2020 as the dataframes from 2020 don’t have data for this party. 
Q4
Q5
- We may not be able to compare the change in the number of registrations for LIB and NLM parties (from primary), and NLM parties (from general) between 2024 and 2020 as the dataframes from 2020 don’t have data for these parties.

*Question 4* Describe the limits of the data in terms of timeframe, what is and is not included and any codes or other information you'd need to find definitions for, and what you've done to better define it.

*Answer 4*
Q1
- The registration breakdown by county doesn’t have the total row of registrations for each party which we have to use to answer questions at the state level. We may need to find the sum ourselves.
- The population data of Maryland is not included here. We have to download the data from the census.
- We are not sure how to make the comparison between the voter registration and the election result data (https://elections.maryland.gov/elections/2024/general_Results/gen_results_2024_1.html) makes sense because the registration is categorized by parties but the result is categorized by candidates which some of them come from the same party.    
Q2
- We’re not sure whether there is population data for each congressional district, so the part of this question about comparing registration with population may not be able to answer. 
Q3
- We’re not sure whether there is population data for each legislative district, so the part of this question about comparing registration with population may not be able to answer. 
- We’re not sure which election result we should use to compare with the registration data at the legislative level. 
Q4
- We’re not sure where or how to locate registration at the precinct level.
Q5
- The population data of each county in Maryland is not included here. We have to download the data from the census.
- We are not sure how to make the comparison between the voter registration and the election result data (https://elections.maryland.gov/elections/2024/general_Results/gen_detail_results_2024_1_1.html) makes sense because the registration is categorized by parties but the result is categorized by candidates which some of them come from the same party.

-------------------------------------------------------------------------------------------------

*Deliverable 3 (week 13):* 

Turn in, via GitHub, a draft of the data notebook you will produce for this project. You should work on this notebook together, but everyone needs to upload a copy to their own GitHub repo and turn in a link for grading. At a minimum, your data notebook should load the data you'll use with this project and have demonstrated work on at least two of the questions you are seeking to answer. The more you have finished, the more feedback I'll be able to give you.

Q1. For the big picture, in *Maryland*,
1.1 How many voters registered for each party in the 2024 primary and general election?
*Answer: In the 2024 general elections, the registration totals were as follows: 2211316 registered for the Democratic Party, 1007384 for the Republican Party, 78234 for other parties, and 907638 were unaffiliated. In the primary elections, 2202830 registered for the Democrats, 994388 for the Republicans, 31418 for other parties, and 451465 were unaffiliated.*
1.1.1 How have these numbers changed compared to those in previous elections?
*Answer: Compared to the previous election (2020, the primary elections in 2024 saw a decline in registrations for the major parties, while other parties and unaffiliated registrants gained significant popularity. Specifically, the Democratic Party lost 1.24%, the Republican Party lost 1.45%, while other parties gained 44.95%, and unaffiliated registrations increased by 40.24%. This trend continued in the general elections, with the Democratic Party experiencing a 2.27% decrease, while the Republican Party saw a smaller drop of 0.82%. Similarly, other parties gained 21.54%, and unaffiliated registrations rose by 18.36%.*
1.2 Compared to the total population of the state, what is the proportion of voters who register for the election?
*Answer: In the primary elections, voter turnout was 59.72% of the total population, while in the general elections, it increased to 68.23%.*
1.3 Compared to the 2024 presidential election result, does the registration pattern match with the result?
*Answer: By comparing the general party registrations in Maryland for 2024 with the presidential election results, it is notable that the number of actual voters for most parties and the unaffiliated group sharply decreased. The exception to this trend is the Republican Party, which gained 19,258 more votes than the number of registrations. In contrast, the Democratic Party saw a significant decline, losing 350,372 voters in the election compared to their registrations.*

```{r}
#Prep dataframe for Q1 focusing on Maryland
#Registration Data
md_primary_24 <- read_csv("data/county/county_primary_24.csv")
md_general_24 <- read_csv("data/county/county_general_24.csv")
md_primary_20 <- read_csv("data/county/county_primary_20.csv")
md_primary_20 <- md_primary_20 |>
  rename(County = 1)
md_general_20 <- read_csv("data/county/county_general_20.csv")
#Population Data
state_population <- get_acs(geography="state", 
                       variables = "B01003_001", 
                       year=2022)
md_population <- state_population |>
  slice(21)
#Presidential Election Result
pres_election_result_url <- "https://elections.maryland.gov/elections/2024/general_Results/gen_results_2024_1.html"
pres_election_result_html <- pres_election_result_url |>
  read_html() |>
  html_table()
pres_election_result <- pres_election_result_html[[1]]
pres_election_result <- pres_election_result |> clean_names()
pres_election_result <- pres_election_result |>
    slice(-c(34, 35))
```
```{r}
#voters registered for each party in 2024 primary election
total_md_primary_24 <- md_primary_24 |> 
    mutate(OTH = OTH + NLM + LIB) |>
  summarise(
    DEM_tot = sum(DEM),
    REP_tot = sum(REP),
    OTH_tot = sum(OTH),
    UNA_tot = sum(UNA)
  )
```
```{r}
#voters registered for each party in 2024 general election
total_md_general_24 <- md_general_24 |> 
  mutate(OTH = OTH + GRN + NLM + LIB) |> 
summarise(
    DEM_tot = sum(DEM),
    REP_tot = sum(REP),
    OTH_tot = sum(OTH),
    UNA_tot = sum(UNA)
  )
```
```{r}
#voters registered for each party in 2020 primary election
total_md_primary_20 <- md_primary_20 |>
  mutate(OTH = OTH + BAR + WCP) |>
  summarise(
    DEM_tot = sum(DEM),
    REP_tot = sum(REP),
    OTH_tot = sum(OTH),
    UNA_tot = sum(UNA)
  )
```
```{r}
#voters registered for each party in 2020 general election
total_md_general_20 <- md_general_20 |>
  mutate(OTH = OTH + BAR + WCP + GRN + LIB) |>
  summarise(
    DEM_tot = sum(DEM),
    REP_tot = sum(REP),
    OTH_tot = sum(OTH),
    UNA_tot = sum(UNA)
  )
```
```{r}
#How have these numbers changed compared to those in previous elections?*
total_md_primary_24 <- total_md_primary_24 |> 
  mutate(Year = 2024)
total_md_general_24 <- total_md_general_24 |> 
  mutate(Year = 2024)
total_md_primary_20 <- total_md_primary_20 |> 
  mutate(Year = 2020)
total_md_general_20 <- total_md_general_20 |> 
  mutate(Year = 2020)
comparison_general <- bind_rows(total_md_general_24, total_md_general_20) |>
select(Year, everything())
comparison_primary <- bind_rows(total_md_primary_24, total_md_primary_20) |>
select(Year, everything())
```
```{r}
#percentage difference between 2024 and 2022 per each party in geenral elections
comparison_general <- comparison_general|>
    mutate(
    DEM_pct_diff = (DEM_tot[1] - DEM_tot[2]) / DEM_tot[2] * 100,
    REP_pct_diff = (REP_tot[1] - REP_tot[2]) / REP_tot[2] * 100,
    OTH_pct_diff = (OTH_tot[1] - OTH_tot[2]) / OTH_tot[2] * 100,
    UNA_pct_diff = (UNA_tot[1] - UNA_tot[2]) / UNA_tot[2] * 100
  )
```
```{r}
#percentage difference between 2024 and 2022 per each party in primary elections
comparison_primary <- comparison_primary|>
    mutate(
    DEM_pct_diff = (DEM_tot[1] - DEM_tot[2]) / DEM_tot[2] * 100,
    REP_pct_diff = (REP_tot[1] - REP_tot[2]) / REP_tot[2] * 100,
    OTH_pct_diff = (OTH_tot[1] - OTH_tot[2]) / OTH_tot[2] * 100,
    UNA_pct_diff = (UNA_tot[1] - UNA_tot[2]) / UNA_tot[2] * 100
  )
```
```{r}
#voters registered for general 2024 in proportion to population
proportion_md_general_24 <- md_general_24 |>
  summarise(
    total_voters = sum(TOTAL),
    MD_pop = "6161707",
    pct_voters = (total_voters / 6161707) *100
  )
```
```{r}
#voters registered for primary 2024 in proportion to population
proportion_md_primary_24 <- md_primary_24 |>
  summarise(
    total_voters = sum(TOTAL),
    MD_pop = "6161707",
    pct_voters = (total_voters / 6161707) *100
  )
```
```{r}
#*1.3 Compared to the 2024 presidential election result, does the registration pattern match with the result?*
party_elections_results <- pres_election_result |>
  mutate(
    total = as.numeric(gsub(",", "", total))  
  ) |>
  group_by(party) |> 
  summarise(
    Total_votes = sum(total)  
  )
party_elections_results <- party_elections_results |> 
  mutate(
    Total_votes = if_else(party == "Other Candidates", 
                          Total_votes + sum(Total_votes[party %in% c("Green", "Libertarian")]), 
                          Total_votes)
  ) |> 
  filter(!party %in% c("Green", "Libertarian"))
```
```{r}
total_md_general_24_transformed <- total_md_general_24 |> 
  pivot_longer(
    cols = c(DEM_tot, REP_tot, OTH_tot, UNA_tot),  
    names_to = "party",                          
    values_to = "Total_votes"                   
  ) |> 
  mutate(
    party = case_when(                           
      party == "DEM_tot" ~ "Democratic",
      party == "REP_tot" ~ "Republican",
      party == "OTH_tot" ~ "Other Candidates",
      party == "UNA_tot" ~ "Unaffiliated",
      TRUE ~ party
    )
  ) |> 
  select(party, Total_votes)
```
```{r}
comparison_general_elections <- full_join(party_elections_results, total_md_general_24_transformed, by = "party") |>
  rename(Total_election = Total_votes.x, Total_general = Total_votes.y) |>
    mutate(
    Difference = Total_election - Total_general 
    )
```

Q2. At the *Congressional District level*
2.1 How many voters registered for each party in the 2024 primary and general election?
*A: For the 2024 Primary, there are 2,202,830 registrations for Democrats, 994,388 registrations for Republicans, 451,465 registrations for Unaffiliated, and 31,418 registrations for other parties. For 2024 General, there are 2,211,316 registrations for Democrats, 1,007,384 registrations for Republicans, 907,638 registrations for Unaffiliated, and 78,234 registrations for other parties.*
2.1.1 How have these numbers changed compared to those in previous elections?
*A: For the 2024 and 2020 Primary, the registration for Democrats decreased by 1.24%, Republicans decreased by 1.45%, Unaffiliated increased by 40.24%, and other parties increased by 44.96%. For 2024 and 2020 General, the registration for Democrats decreased by 2.28%, Republicans decreased by 0.83%, Unaffiliated increased by 18.37%, and other parties increased by 21.55%.*
2.2 Are there any patterns for voter registration from every district?
*A: For both Primary and General 2024, in most districts, there are registrations for Democrats more than registration for Republicans.*
2.2.1 Are there any districts that don't fall into this pattern?
*A: There is only District 1 where registration for Democrats is less than that of Republicans.*
2.2.2 How are those districts different from others in terms of location, demography, economy, etc.?
*A:.....* 
2.3 Compared to the total population of each district, what is the proportion of voters who register for the election? 
*A: For both Primary and General, District 5 has the most proportion of registered voters which are 69.26% for Primary and 74.12% for General. In the other hand, District 4 has the least proportion of registered voters which are 52.02% for Primary and 58.34% for General* 
2.3.1 What are the differences between district having large and small proportions of voter registration?
*A:.....* 
2.4 Compared to the representative race result, does the registration pattern match with the result?
*A: Yes, the General registration pattern in every district is match with the pattern of the representative race result. There are more percentage of people who registered and votes for Democrats than those for Republicans in District 2 to 8, and on the contrary, there are more percentage of people who registered and votes for Republicans than those for Democrats in District 1.* 

```{r}
#Prep dataframe for Q2 focusing on congressional district level

#Registration Data
congress_primary_24 <- read_csv("data/congress_level/congress_primary_24.csv")
congress_general_24 <- read_csv("data/congress_level/congress_general_24.csv")
congress_primary_20 <- read_csv("data/congress_level/congress_primary_20.csv")
congress_primary_20 <- congress_primary_20 |>
  rename(District = 1) |>
  rename(county = 2) |>
  rename(BAR = 3) |>
  rename(DEM = 4) |>
  rename(OTH = 5) |>
  rename(REP = 6) |>
  rename(UNA = 7) |>
  rename(WCP = 8) |>
  rename(total = 9) 
congress_general_20 <- read_csv("data/congress_level/congress_general_20.csv")

#Population Data
congress_population <- get_acs(geography="congressional district", 
                       variables = "B01003_001", state = "MD", 
                       year=2022)

#Representative Election Result
representative_election_result_url <- "https://elections.maryland.gov/elections/2024/general_Results/gen_results_2024_3.html"
representative_election_result_html <- representative_election_result_url |>
  read_html() |>
  html_table()
representative_election_result1 <- representative_election_result_html[[1]]
representative_election_result2 <- representative_election_result_html[[2]]
representative_election_result3 <- representative_election_result_html[[3]]
representative_election_result4 <- representative_election_result_html[[4]]
representative_election_result5 <- representative_election_result_html[[5]]
representative_election_result6 <- representative_election_result_html[[6]]
representative_election_result7 <- representative_election_result_html[[7]]
representative_election_result8 <- representative_election_result_html[[8]]
```
```{r}
#Primary
#2024
count_congress_primary_24 <- congress_primary_24 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    LIB = sum(LIB),
    OTH = sum(OTH),
    UNA = sum(UNA),
    NLM = sum(NLM)
  ) |>
  mutate(Other = LIB+OTH+NLM) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM24 = 1, REP24 =2, UNA24 = 3, Other24 = 4)
#2020
count_congress_primary_20 <- congress_primary_20 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    OTH = sum(OTH),
    UNA = sum(UNA),
    BAR = sum(BAR),
    WCP = sum(WCP)
  ) |>
  mutate(Other = OTH+BAR+WCP) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM20 = 1, REP20 =2, UNA20 = 3, Other20 = 4)
#Find the change
compare_congress_primary_24_20 <- bind_cols(count_congress_primary_24, count_congress_primary_20) |>
  mutate(DEM_change = ((DEM24-DEM20)/DEM20)*100) |>
  mutate(REP_change = ((REP24-REP20)/REP20)*100) |>
  mutate(UNA_change = ((UNA24-UNA20)/UNA20)*100) |>
  mutate(Other_change = ((Other24-Other20)/Other20)*100)

#General
#2024
count_congress_general_24 <- congress_general_24 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    LIB = sum(LIB),
    OTH = sum(OTH),
    UNA = sum(UNA),
    NLM = sum(NLM),
    GRN = sum(GRN)
  ) |>
  mutate(Other = LIB+OTH+NLM+GRN) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM24 = 1, REP24 =2, UNA24 = 3, Other24 = 4)
#2020
count_congress_general_20 <- congress_general_20 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    BAR = sum(BAR),
    GRN = sum(GRN),
    LIB = sum(LIB),
    WCP = sum(WCP),
    OTH = sum(OTH),
    UNA = sum(UNA)
  ) |>
  mutate(Other = OTH+BAR+GRN+LIB+WCP) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM20 = 1, REP20 =2, UNA20 = 3, Other20 = 4)
#Find the change
compare_congress_general_24_20 <- bind_cols(count_congress_general_24, count_congress_general_20) |>
  mutate(DEM_change = ((DEM24-DEM20)/DEM20)*100) |>
  mutate(REP_change = ((REP24-REP20)/REP20)*100) |>
  mutate(UNA_change = ((UNA24-UNA20)/UNA20)*100) |>
  mutate(Other_change = ((Other24-Other20)/Other20)*100)

#Find Pattern
#Primary24
pattern_congress_primary_24 <- congress_primary_24 |>
  mutate(Other = LIB+OTH+NLM) |>
  select(District, DEM, REP, UNA, Other, TOTAL) |>
  mutate(DEM_per = (DEM/TOTAL)*100) |>
  mutate(REP_per = (REP/TOTAL)*100) |>
  mutate(UNA_per = (UNA/TOTAL)*100) |>
  mutate(Other_per = (Other/TOTAL)*100)
#General24
pattern_congress_general_24 <- congress_general_24 |>
  mutate(Other = LIB+OTH+NLM+GRN) |>
  select(District, DEM, REP, UNA, Other, TOTAL) |>
  mutate(DEM_per = (DEM/TOTAL)*100) |>
  mutate(REP_per = (REP/TOTAL)*100) |>
  mutate(UNA_per = (UNA/TOTAL)*100) |>
  mutate(Other_per = (Other/TOTAL)*100)

#Compare to Population
#Primary24
total_congress_primary_24 <- congress_primary_24 |>
  select(District, TOTAL)
cleaned_congress_population <- congress_population |>
  rename(population = 4) |>
  select(4)
regis_population_congress_primary_24 <- bind_cols(total_congress_primary_24, cleaned_congress_population) |>
  mutate(Percent = (TOTAL/population)*100)
#General24
total_congress_general_24 <- congress_general_24 |>
  select(District, TOTAL)
regis_population_congress_general_24 <- bind_cols(total_congress_general_24, cleaned_congress_population) |>
  mutate(Percent = (TOTAL/population)*100)

#Compare to the result
#Clean the result
representative_election_result1 <- representative_election_result1 |>
  slice(-5) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, x = 4, y = 5) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(x = as.numeric(gsub("%","", x))) |>
  mutate(y = as.numeric(gsub("%","", y))) |>
  mutate(Other_percent_result = x+y) |>
  select(1,2,3,6)
representative_election_result2 <- representative_election_result2 |>
  slice(-5) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, x = 4, y = 5) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(x = as.numeric(gsub("%","", x))) |>
  mutate(y = as.numeric(gsub("%","", y))) |>
  mutate(Other_percent_result = x+y) |>
  select(1,2,3,6)
representative_election_result3 <- representative_election_result3 |>
  slice(-5) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, x = 4, y = 5) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(x = as.numeric(gsub("%","", x))) |>
  mutate(y = as.numeric(gsub("%","", y))) |>
  mutate(Other_percent_result = x+y) |>
  select(1,2,3,6)
representative_election_result4 <- representative_election_result4 |>
  slice(-4) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, Other_percent_result = 4) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(Other_percent_result = as.numeric(gsub("%","", Other_percent_result)))
representative_election_result5 <- representative_election_result5 |>
  slice(-4) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, Other_percent_result = 4) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(Other_percent_result = as.numeric(gsub("%","", Other_percent_result)))
representative_election_result6 <- representative_election_result6 |>
  slice(-4) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, Other_percent_result = 4) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(Other_percent_result = as.numeric(gsub("%","", Other_percent_result)))
representative_election_result7 <- representative_election_result7 |>
  slice(-5) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, x = 4, y = 5) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(x = as.numeric(gsub("%","", x))) |>
  mutate(y = as.numeric(gsub("%","", y))) |>
  mutate(Other_percent_result = x+y) |>
  select(1,2,3,6)
representative_election_result8 <- representative_election_result8 |>
  slice(-5) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, x = 4, y = 5) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(x = as.numeric(gsub("%","", x))) |>
  mutate(y = as.numeric(gsub("%","", y))) |>
  mutate(Other_percent_result = x+y) |>
  select(1,2,3,6)
representative_election_result <- bind_rows(list(representative_election_result1, representative_election_result2, representative_election_result3, representative_election_result4, representative_election_result5, representative_election_result6, representative_election_result7, representative_election_result8)) |>
  select(2,3,4)
#Merge with the registration
congress_general_regis_result <- bind_cols(pattern_congress_general_24, representative_election_result) |>
  select(-2,-3,-4,-5,-6)
```

-------------------------------------------------------------------------------------------------

*Deliverable 4* 

- Shows a draft of data notebook in which you demonstrate substantial progress on the questions you set out to answer at the beginning of the project. This means you should have started on each of them, but you may not have finished.

Q1. For the big picture, in *Maryland*,
1.1 How many voters registered for each party in the 2024 primary and general election?
*Answer: In the 2024 general elections, the registration totals were as follows: 2211316 registered for the Democratic Party, 1007384 for the Republican Party, 78234 for other parties, and 907638 were unaffiliated. In the primary elections, 2202830 registered for the Democrats, 994388 for the Republicans, 31418 for other parties, and 451465 were unaffiliated.*
1.1.1 How have these numbers changed compared to those in previous elections?
*Answer: Compared to the previous election (2020, the primary elections in 2024 saw a decline in registrations for the major parties, while other parties and unaffiliated registrants gained significant popularity. Specifically, the Democratic Party lost 1.24%, the Republican Party lost 1.45%, while other parties gained 44.95%, and unaffiliated registrations increased by 40.24%. This trend continued in the general elections, with the Democratic Party experiencing a 2.27% decrease, while the Republican Party saw a smaller drop of 0.82%. Similarly, other parties gained 21.54%, and unaffiliated registrations rose by 18.36%.*
1.2 Compared to the total population of the state, what is the proportion of voters who register for the election?
*Answer: In the primary elections, voter turnout was 59.72% of the total population, while in the general elections, it increased to 68.23%.*
1.3 Compared to the 2024 presidential election result, does the registration pattern match with the result?
*Answer: By comparing the general party registrations in Maryland for 2024 with the presidential election results, it is notable that the number of actual voters for most parties and the unaffiliated group sharply decreased. The exception to this trend is the Republican Party, which gained 19,258 more votes than the number of registrations. In contrast, the Democratic Party saw a significant decline, losing 350,372 voters in the election compared to their registrations.*

```{r}
#Prep dataframe for Q1 focusing on Maryland
#Registration Data
md_primary_24 <- read_csv("data/county/county_primary_24.csv")
md_general_24 <- read_csv("data/county/county_general_24.csv")
md_primary_20 <- read_csv("data/county/county_primary_20.csv")
md_primary_20 <- md_primary_20 |>
  rename(County = 1)
md_general_20 <- read_csv("data/county/county_general_20.csv")
#Population Data
state_population <- get_acs(geography="state", 
                       variables = "B01003_001", 
                       year=2022)
md_population <- state_population |>
  slice(21)
#Presidential Election Result
pres_election_result_url <- "https://elections.maryland.gov/elections/2024/general_Results/gen_results_2024_1.html"
pres_election_result_html <- pres_election_result_url |>
  read_html() |>
  html_table()
pres_election_result <- pres_election_result_html[[1]]
pres_election_result <- pres_election_result |> clean_names()
pres_election_result <- pres_election_result |>
    slice(-c(34, 35))
```
```{r}
#voters registered for each party in 2024 primary election
total_md_primary_24 <- md_primary_24 |> 
    mutate(OTH = OTH + NLM + LIB) |>
  summarise(
    DEM_tot = sum(DEM),
    REP_tot = sum(REP),
    OTH_tot = sum(OTH),
    UNA_tot = sum(UNA)
  )
```
```{r}
#voters registered for each party in 2024 general election
total_md_general_24 <- md_general_24 |> 
  mutate(OTH = OTH + GRN + NLM + LIB) |> 
summarise(
    DEM_tot = sum(DEM),
    REP_tot = sum(REP),
    OTH_tot = sum(OTH),
    UNA_tot = sum(UNA)
  )
```
```{r}
#voters registered for each party in 2020 primary election
total_md_primary_20 <- md_primary_20 |>
  mutate(OTH = OTH + BAR + WCP) |>
  summarise(
    DEM_tot = sum(DEM),
    REP_tot = sum(REP),
    OTH_tot = sum(OTH),
    UNA_tot = sum(UNA)
  )
```
```{r}
#voters registered for each party in 2020 general election
total_md_general_20 <- md_general_20 |>
  mutate(OTH = OTH + BAR + WCP + GRN + LIB) |>
  summarise(
    DEM_tot = sum(DEM),
    REP_tot = sum(REP),
    OTH_tot = sum(OTH),
    UNA_tot = sum(UNA)
  )
```
```{r}
#How have these numbers changed compared to those in previous elections?*
total_md_primary_24 <- total_md_primary_24 |> 
  mutate(Year = 2024)
total_md_general_24 <- total_md_general_24 |> 
  mutate(Year = 2024)
total_md_primary_20 <- total_md_primary_20 |> 
  mutate(Year = 2020)
total_md_general_20 <- total_md_general_20 |> 
  mutate(Year = 2020)
comparison_general <- bind_rows(total_md_general_24, total_md_general_20) |>
select(Year, everything())
comparison_primary <- bind_rows(total_md_primary_24, total_md_primary_20) |>
select(Year, everything())
```
```{r}
#percentage difference between 2024 and 2022 per each party in geenral elections
comparison_general <- comparison_general|>
    mutate(
    DEM_pct_diff = (DEM_tot[1] - DEM_tot[2]) / DEM_tot[2] * 100,
    REP_pct_diff = (REP_tot[1] - REP_tot[2]) / REP_tot[2] * 100,
    OTH_pct_diff = (OTH_tot[1] - OTH_tot[2]) / OTH_tot[2] * 100,
    UNA_pct_diff = (UNA_tot[1] - UNA_tot[2]) / UNA_tot[2] * 100
  )
```
```{r}
#percentage difference between 2024 and 2022 per each party in primary elections
comparison_primary <- comparison_primary|>
    mutate(
    DEM_pct_diff = (DEM_tot[1] - DEM_tot[2]) / DEM_tot[2] * 100,
    REP_pct_diff = (REP_tot[1] - REP_tot[2]) / REP_tot[2] * 100,
    OTH_pct_diff = (OTH_tot[1] - OTH_tot[2]) / OTH_tot[2] * 100,
    UNA_pct_diff = (UNA_tot[1] - UNA_tot[2]) / UNA_tot[2] * 100
  )
```
```{r}
#voters registered for general 2024 in proportion to population
proportion_md_general_24 <- md_general_24 |>
  summarise(
    total_voters = sum(TOTAL),
    MD_pop = "6161707",
    pct_voters = (total_voters / 6161707) *100
  )
```
```{r}
#voters registered for primary 2024 in proportion to population
proportion_md_primary_24 <- md_primary_24 |>
  summarise(
    total_voters = sum(TOTAL),
    MD_pop = "6161707",
    pct_voters = (total_voters / 6161707) *100
  )
```
```{r}
#*1.3 Compared to the 2024 presidential election result, does the registration pattern match with the result?*
party_elections_results <- pres_election_result |>
  mutate(
    total = as.numeric(gsub(",", "", total))  
  ) |>
  group_by(party) |> 
  summarise(
    Total_votes = sum(total)  
  )
party_elections_results <- party_elections_results |> 
  mutate(
    Total_votes = if_else(party == "Other Candidates", 
                          Total_votes + sum(Total_votes[party %in% c("Green", "Libertarian")]), 
                          Total_votes)
  ) |> 
  filter(!party %in% c("Green", "Libertarian"))
```
```{r}
total_md_general_24_transformed <- total_md_general_24 |> 
  pivot_longer(
    cols = c(DEM_tot, REP_tot, OTH_tot, UNA_tot),  
    names_to = "party",                          
    values_to = "Total_votes"                   
  ) |> 
  mutate(
    party = case_when(                           
      party == "DEM_tot" ~ "Democratic",
      party == "REP_tot" ~ "Republican",
      party == "OTH_tot" ~ "Other Candidates",
      party == "UNA_tot" ~ "Unaffiliated",
      TRUE ~ party
    )
  ) |> 
  select(party, Total_votes)
```
```{r}
comparison_general_elections <- full_join(party_elections_results, total_md_general_24_transformed, by = "party") |>
  rename(Total_election = Total_votes.x, Total_general = Total_votes.y) |>
    mutate(
    Difference = Total_election - Total_general 
    )
```

Q2. At the *Congressional District level*
2.1 How many voters registered for each party in the 2024 primary and general election?
*A: For the 2024 Primary, there are 2,202,830 registrations for Democrats, 994,388 registrations for Republicans, 451,465 registrations for Unaffiliated, and 31,418 registrations for other parties. For 2024 General, there are 2,211,316 registrations for Democrats, 1,007,384 registrations for Republicans, 907,638 registrations for Unaffiliated, and 78,234 registrations for other parties.*
2.1.1 How have these numbers changed compared to those in previous elections?
*A: For the 2024 and 2020 Primary, the registration for Democrats decreased by 1.24%, Republicans decreased by 1.45%, Unaffiliated increased by 40.24%, and other parties increased by 44.96%. For 2024 and 2020 General, the registration for Democrats decreased by 2.28%, Republicans decreased by 0.83%, Unaffiliated increased by 18.37%, and other parties increased by 21.55%.*
2.2 Are there any patterns for voter registration from every district?
*A: For both Primary and General 2024, in most districts, there are registrations for Democrats more than registration for Republicans.*
2.2.1 Are there any districts that don't fall into this pattern?
*A: There is only District 1 where registration for Democrats is less than that of Republicans.*
2.2.2 How are those districts different from others in terms of location, demography, economy, etc.?
*A: It is all the area located on the east side of Chesapeake Bay.* 
2.3 Compared to the total population of each district, what is the proportion of voters who register for the election? 
*A: For both Primary and General, District 5 has the most proportion of registered voters which are 69.26% for Primary and 74.12% for General. In the other hand, District 4 has the least proportion of registered voters which are 52.02% for Primary and 58.34% for General* 
2.3.1 What are the differences between district having large and small proportions of voter registration?
*A: There is no significant difference.* 
2.4 Compared to the representative race result, does the registration pattern match with the result?
*A: Yes, the General registration pattern in every district is match with the pattern of the representative race result. There are more percentage of people who registered and votes for Democrats than those for Republicans in District 2 to 8, and on the contrary, there are more percentage of people who registered and votes for Republicans than those for Democrats in District 1.* 

```{r}
#Prep dataframe for Q2 focusing on congressional district level

#Registration Data
congress_primary_24 <- read_csv("data/congress_level/congress_primary_24.csv")
congress_general_24 <- read_csv("data/congress_level/congress_general_24.csv")
congress_primary_20 <- read_csv("data/congress_level/congress_primary_20.csv")
congress_primary_20 <- congress_primary_20 |>
  rename(District = 1) |>
  rename(county = 2) |>
  rename(BAR = 3) |>
  rename(DEM = 4) |>
  rename(OTH = 5) |>
  rename(REP = 6) |>
  rename(UNA = 7) |>
  rename(WCP = 8) |>
  rename(total = 9) 
congress_general_20 <- read_csv("data/congress_level/congress_general_20.csv")

#Population Data
congress_population <- get_acs(geography="congressional district", 
                       variables = "B01003_001", state = "MD", 
                       year=2022)

#Representative Election Result
representative_election_result_url <- "https://elections.maryland.gov/elections/2024/general_Results/gen_results_2024_3.html"
representative_election_result_html <- representative_election_result_url |>
  read_html() |>
  html_table()
representative_election_result1 <- representative_election_result_html[[1]]
representative_election_result2 <- representative_election_result_html[[2]]
representative_election_result3 <- representative_election_result_html[[3]]
representative_election_result4 <- representative_election_result_html[[4]]
representative_election_result5 <- representative_election_result_html[[5]]
representative_election_result6 <- representative_election_result_html[[6]]
representative_election_result7 <- representative_election_result_html[[7]]
representative_election_result8 <- representative_election_result_html[[8]]
```
```{r}
#Primary
#2024
count_congress_primary_24 <- congress_primary_24 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    LIB = sum(LIB),
    OTH = sum(OTH),
    UNA = sum(UNA),
    NLM = sum(NLM)
  ) |>
  mutate(Other = LIB+OTH+NLM) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM24 = 1, REP24 =2, UNA24 = 3, Other24 = 4)
#2020
count_congress_primary_20 <- congress_primary_20 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    OTH = sum(OTH),
    UNA = sum(UNA),
    BAR = sum(BAR),
    WCP = sum(WCP)
  ) |>
  mutate(Other = OTH+BAR+WCP) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM20 = 1, REP20 =2, UNA20 = 3, Other20 = 4)
#Find the change
compare_congress_primary_24_20 <- bind_cols(count_congress_primary_24, count_congress_primary_20) |>
  mutate(DEM_change = ((DEM24-DEM20)/DEM20)*100) |>
  mutate(REP_change = ((REP24-REP20)/REP20)*100) |>
  mutate(UNA_change = ((UNA24-UNA20)/UNA20)*100) |>
  mutate(Other_change = ((Other24-Other20)/Other20)*100)

#General
#2024
count_congress_general_24 <- congress_general_24 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    LIB = sum(LIB),
    OTH = sum(OTH),
    UNA = sum(UNA),
    NLM = sum(NLM),
    GRN = sum(GRN)
  ) |>
  mutate(Other = LIB+OTH+NLM+GRN) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM24 = 1, REP24 =2, UNA24 = 3, Other24 = 4)
#2020
count_congress_general_20 <- congress_general_20 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    BAR = sum(BAR),
    GRN = sum(GRN),
    LIB = sum(LIB),
    WCP = sum(WCP),
    OTH = sum(OTH),
    UNA = sum(UNA)
  ) |>
  mutate(Other = OTH+BAR+GRN+LIB+WCP) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM20 = 1, REP20 =2, UNA20 = 3, Other20 = 4)
#Find the change
compare_congress_general_24_20 <- bind_cols(count_congress_general_24, count_congress_general_20) |>
  mutate(DEM_change = ((DEM24-DEM20)/DEM20)*100) |>
  mutate(REP_change = ((REP24-REP20)/REP20)*100) |>
  mutate(UNA_change = ((UNA24-UNA20)/UNA20)*100) |>
  mutate(Other_change = ((Other24-Other20)/Other20)*100)

#Find Pattern
#Primary24
pattern_congress_primary_24 <- congress_primary_24 |>
  mutate(Other = LIB+OTH+NLM) |>
  select(District, DEM, REP, UNA, Other, TOTAL) |>
  mutate(DEM_per = (DEM/TOTAL)*100) |>
  mutate(REP_per = (REP/TOTAL)*100) |>
  mutate(UNA_per = (UNA/TOTAL)*100) |>
  mutate(Other_per = (Other/TOTAL)*100)
#General24
pattern_congress_general_24 <- congress_general_24 |>
  mutate(Other = LIB+OTH+NLM+GRN) |>
  select(District, DEM, REP, UNA, Other, TOTAL) |>
  mutate(DEM_per = (DEM/TOTAL)*100) |>
  mutate(REP_per = (REP/TOTAL)*100) |>
  mutate(UNA_per = (UNA/TOTAL)*100) |>
  mutate(Other_per = (Other/TOTAL)*100)

#Compare to Population
#Primary24
total_congress_primary_24 <- congress_primary_24 |>
  select(District, TOTAL)
cleaned_congress_population <- congress_population |>
  rename(population = 4) |>
  select(4)
regis_population_congress_primary_24 <- bind_cols(total_congress_primary_24, cleaned_congress_population) |>
  mutate(Percent = (TOTAL/population)*100)
#General24
total_congress_general_24 <- congress_general_24 |>
  select(District, TOTAL)
regis_population_congress_general_24 <- bind_cols(total_congress_general_24, cleaned_congress_population) |>
  mutate(Percent = (TOTAL/population)*100)

#Compare to the result
#Clean the result
representative_election_result1 <- representative_election_result1 |>
  slice(-5) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, x = 4, y = 5) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(x = as.numeric(gsub("%","", x))) |>
  mutate(y = as.numeric(gsub("%","", y))) |>
  mutate(Other_percent_result = x+y) |>
  select(1,2,3,6)
representative_election_result2 <- representative_election_result2 |>
  slice(-5) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, x = 4, y = 5) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(x = as.numeric(gsub("%","", x))) |>
  mutate(y = as.numeric(gsub("%","", y))) |>
  mutate(Other_percent_result = x+y) |>
  select(1,2,3,6)
representative_election_result3 <- representative_election_result3 |>
  slice(-5) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, x = 4, y = 5) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(x = as.numeric(gsub("%","", x))) |>
  mutate(y = as.numeric(gsub("%","", y))) |>
  mutate(Other_percent_result = x+y) |>
  select(1,2,3,6)
representative_election_result4 <- representative_election_result4 |>
  slice(-4) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, Other_percent_result = 4) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(Other_percent_result = as.numeric(gsub("%","", Other_percent_result)))
representative_election_result5 <- representative_election_result5 |>
  slice(-4) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, Other_percent_result = 4) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(Other_percent_result = as.numeric(gsub("%","", Other_percent_result)))
representative_election_result6 <- representative_election_result6 |>
  slice(-4) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, Other_percent_result = 4) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(Other_percent_result = as.numeric(gsub("%","", Other_percent_result)))
representative_election_result7 <- representative_election_result7 |>
  slice(-5) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, x = 4, y = 5) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(x = as.numeric(gsub("%","", x))) |>
  mutate(y = as.numeric(gsub("%","", y))) |>
  mutate(Other_percent_result = x+y) |>
  select(1,2,3,6)
representative_election_result8 <- representative_election_result8 |>
  slice(-5) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, x = 4, y = 5) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(x = as.numeric(gsub("%","", x))) |>
  mutate(y = as.numeric(gsub("%","", y))) |>
  mutate(Other_percent_result = x+y) |>
  select(1,2,3,6)
representative_election_result <- bind_rows(list(representative_election_result1, representative_election_result2, representative_election_result3, representative_election_result4, representative_election_result5, representative_election_result6, representative_election_result7, representative_election_result8)) |>
  select(2,3,4)
#Merge with the registration
congress_general_regis_result <- bind_cols(pattern_congress_general_24, representative_election_result) |>
  select(-2,-3,-4,-5,-6)
```

Q3. At the *Legislative Districts level*
3.1: Which legislative districts had the closest margins between Democrat and Republican parties?
3.2: Do unaffiliated voters constitute a decent portion of the election?

```{r}
#Prep dataframe for Q3 focusing on legislative district level

#Registration Data
legis_primary_24 <- read_csv("data/legis_level/legis_primary_24.csv")
legis_general_24 <- read_csv("data/legis_level/legis_general_24.csv")
legis_primary_20 <- read_csv("data/legis_level/legis_primary_20.csv")
legis_general_20 <- read_csv("data/legis_level/legis_general_20.csv")

#Population Data
legis_population <- get_acs(geography="state legislative district (lower chamber)", 
                       variables = "B01003_001", state = "MD", 
                       year=2022)

#Presidential Election Result 
#Download Precinct Result and Code
precint_result <- read_csv("data/legis_level/result/result.csv")
legis_code <- read_csv("data/legis_level/result/code.csv")
#Prep result dataframe
precint_result <- precint_result |>
  filter(Contest == "President and Vice President of the United States")
precint_result <- precint_result |>
  select(2,4,6,7) |>
  clean_names() |>
  mutate(precinct_name = as.numeric(gsub("-","", precinct_name)))
#Prep code dataframe
legis_code <- legis_code |>
  select(3,4,10) |>
  filter(!is.na(ED_PRECINCT)) |>
  filter(!is.na(LEGISLATIVE_SUB_DISTRICT_CODE))
#Join result dataframe with code dataframe to generate legislative district results 
legis_result_draft <- precint_result |> 
  left_join(legis_code, join_by(jurisdiction==COUNTY_NAME, precinct_name==ED_PRECINCT)) |>
  filter(!is.na(LEGISLATIVE_SUB_DISTRICT_CODE)) |>
  group_by(LEGISLATIVE_SUB_DISTRICT_CODE, party) |>
  summarise(
    total_votes = sum(total_votes))
#Reshape dataframe
legis_result_DEM <- legis_result_draft |>
  filter(party == "DEM") |>
  select(1,3) |>
  rename(DEM_result = 2)
legis_result_GRN <- legis_result_draft |>
  filter(party == "GRN") |>
  select(3) |>
  rename(GRN_result = 2) 
legis_result_LIB <- legis_result_draft |>
  filter(party == "LIB") |>
  select(3) |>
  rename(LIB_result = 2)
legis_result_NON <- legis_result_draft |>
  filter(party == "NON") |>
  select(3) |>
  rename(NON_result = 2)
legis_result_REP <- legis_result_draft |>
  filter(party == "REP") |>
  select(3) |>
  rename(REP_result = 2)
legis_result_UNF <- legis_result_draft |>
  filter(party == "UNF") |>
  select(3) |>
  rename(UNF_result = 2)
legis_result <- bind_cols(legis_result_DEM, legis_result_GRN, legis_result_LIB, legis_result_NON, legis_result_REP, legis_result_UNF) |>
  select(-3,-5,-7,-9,-11) |>
  rename(legislative_district_code = 1)

```
```{r}
legis_result <- legis_result |>
  mutate(dem_rep_difference = DEM_result - REP_result)|>
  mutate(total_votes = DEM_result + GRN_result + LIB_result + NON_result + REP_result + UNF_result)|>
  arrange(desc(dem_rep_difference))

legis_result <- legis_result |>
  mutate(
    party_dominance = case_when(
      dem_rep_difference > 0 ~ "Democrat Dominant",
      dem_rep_difference < 0 ~ "Republican Dominant",
      TRUE ~ "No Clear Winner" # In case of a tie or no votes
    )
  )
dominance_counts <- legis_result |>
  count(party_dominance)
print(dominance_counts)

closest_races <- legis_result |>
  select(legislative_district_code, DEM_result, REP_result) |>
  mutate(margin = abs(DEM_result - REP_result)) |>
  arrange(margin) |>
  head(10) 

legis_result <- legis_result |>
  mutate(
    unf_percentage = (UNF_result / total_votes) * 100 
  )


legis_result_sorted <- legis_result |>
  arrange(desc(unf_percentage))


head(legis_result_sorted, 10)
```

Q4. At the *Precinct level*
4.1 How many voters registered for each party in the 2024 primary and general election?
*A: Cannot answer as we don't have registration data for precinct level.*
4.1.1 How have these numbers changed compared to those in previous elections?
*A: Cannot answer as we don't have registration data for precinct level.*
4.2 Are there any patterns for voter registration from every precinct?
*A: Cannot answer as we don't have registration data for precinct level.*
4.2.1 Are there any precincts that don't fall into this pattern?
*A: Cannot answer as we don't have registration data for precinct level.*
4.2.2 How are those precincts different from others in terms of location, demography, economy, etc.?
*A: Cannot answer as we don't have registration data for precinct level.*
4.3 Compared to the total population of each precinct, what is the proportion of voters who register for the election?
*A: Cannot answer as we don't have registration data and population data for precinct level.*
4.3.1 What are the differences between precincts having large and small proportions of voter registration?
*A: Cannot answer as we don't have registration data and population data for precinct level.*
4.4 Compared to the presidential election result, does the registration pattern match with the result?
*A: We cannot compare the registration data with the result, but we can indicate some interesting findings about the election results pattern, for instance, in Allegany County where Republican won in most areas (29 precincts), there is 1 precinct that Democrat won; on the contrary, in Montgomery County, where Democrat won in most areas (255 precincts), there is 2 precincts that Republican won.*    

```{r}
#Prep dataframe for Q4 focusing on legislative precinct level

#Presidential Result Data
precinct_result <- read_csv("data/precinct/precinct result.csv")
precinct_result_DEM <- precinct_result |>
  select(1, 2, 3, 4, 6, 7) |>
  filter(Contest == "President and Vice President of the United States") |>
  select(-1) |>
  filter(Party == "DEM") |>
  rename(total_votes_DEM = 5) |>
  select(-4)
precinct_result_REP <- precinct_result |>
  select(1, 2, 3, 4, 6, 7) |>
  filter(Contest == "President and Vice President of the United States") |>
  select(-1) |>
  filter(Party == "REP") |>
  rename(total_votes_REP = 5) |>
  select(-1,-2,-3,-4)
precinct_result_LIB <- precinct_result |>
  select(1, 2, 3, 4, 6, 7) |>
  filter(Contest == "President and Vice President of the United States") |>
  select(-1) |>
  filter(Party == "LIB") |>
  rename(total_votes_LIB = 5) |>
  select(-1,-2,-3,-4)
precinct_result_GRN <- precinct_result |>
  select(1, 2, 3, 4, 6, 7) |>
  filter(Contest == "President and Vice President of the United States") |>
  select(-1) |>
  filter(Party == "GRN") |>
  rename(total_votes_GRN = 5) |>
  select(-1,-2,-3,-4)
precinct_result_UNF <- precinct_result |>
  select(1, 2, 3, 4, 6, 7) |>
  filter(Contest == "President and Vice President of the United States") |>
  select(-1) |>
  filter(Party == "UNF") |>
  rename(total_votes_UNF = 5) |>
  select(-1,-2,-3,-4)
precinct_result_NON <- precinct_result |>
  select(1, 2, 3, 4, 6, 7) |>
  filter(Contest == "President and Vice President of the United States") |>
  select(-1) |>
  filter(Party == "NON") |>
  rename(total_votes_NON = 5) |>
  select(-1,-2,-3,-4)
precinct_result_cleaned <- bind_cols(precinct_result_DEM, precinct_result_REP, precinct_result_LIB, precinct_result_GRN, precinct_result_UNF, precinct_result_NON)
precinct_result_final <- precinct_result_cleaned |>
  mutate(total_votes = total_votes_DEM + total_votes_REP + total_votes_LIB + total_votes_GRN + total_votes_UNF + total_votes_NON)
```
```{r}
#Find the pattern for the results
precinct_result_pattern <- precinct_result_final |>
  mutate(total_votes_others = total_votes_LIB + total_votes_GRN + total_votes_NON) |>
  select(1,2,3,4,5,11,8,10,) |>
  filter(total_votes != 0) |>
  mutate(DEM_percent = (total_votes_DEM/total_votes)*100) |>
  mutate(REP_percent = (total_votes_REP/total_votes)*100) |>
  mutate(others_percent = (total_votes_others/total_votes)*100) |>
  mutate(UNF_percent = (total_votes_UNF/total_votes)*100) |>
  mutate(DEM_REP = DEM_percent - REP_percent) |>
  select(1,2,3,13) |>
  filter(DEM_REP != 0)

#Check each county
precinct_result_pattern_compare <- precinct_result_pattern |>
  mutate(
    D = case_when(
      DEM_REP > 0 ~ 1,
      DEM_REP < 0 ~ 0
      ),
    R = case_when(
      DEM_REP > 0 ~ 0,
      DEM_REP < 0 ~ 1
      )
  ) |>
  group_by(Jurisdiction)|>
  summarise(
    county_count = n(),
    More_DEM = sum(D),
    More_REP = sum(R)
  ) |>
  mutate(More_DEM_percent = (More_DEM/county_count)*100) |>
  mutate(More_REP_percent = (More_REP/county_count)*100)
```

Q5. At the *County level*
5.1 How many voters registered for each party in the 2024 primary and general election?
*A: For the 2024 Primary, there are 2,202,830 registrations for Democrats, 994,388 registrations for Republicans, 451,465 registrations for Unaffiliated, and 31,418 registrations for other parties. For 2024 General, there are 2,211,316 registrations for Democrats, 1,007,384 registrations for Republicans, 907,638 registrations for Unaffiliated, and 78,234 registrations for other parties.*
5.1.1 How have these numbers changed compared to those in previous elections?
*A: For the 2024 and 2020 Primary, the registration for Democrats decreased by 1.24%, Republicans decreased by 1.45%, Unaffiliated increased by 40.24%, and other parties increased by 44.96%. For 2024 and 2020 General, the registration for Democrats decreased by 2.26%, Republicans decreased by 0.83%, Unaffiliated increased by 18.37%, and other parties increased by 21.53%.*
5.2 Are there any patterns for voter registration from every county?
*For both Primary and General in 2024, from 24 counties, almost half of them had people who registered for Democrats more than those for Republicans while the other half were the opposite. And in most counties, there were registrations for Republicans more than those for Unaffiliated.*  
5.2.1 Are there any counties that don't fall into this pattern? 
*5 counties had people who registered for Unaffiliated more than those for Republicans.*
5.2.2 How are those counties different from others in terms of location, demography, economy, etc.?
*Most of them which are Montgomery, Prince George's, Charles, and Howard are located in the center part of Maryland.*
5.3 Compared to the total population of each county, what is the proportion of voters who register for the election?
*Every county had people registering to vote more than 50% of the population.*
5.3.1 What are the differences between counties having large and small proportions of voter registration?
*There are no significant differences.* 
5.4 Compared to the 2024 presidential election result, does the registration pattern match with the result?
*In most counties, the election results had the same pattern as registrations.*

```{r}
#Prep dataframe for Q5 focusing on county level

#Registration Data
county_primary_24 <- read_csv("data/county/county_primary_24.csv")
county_general_24 <- read_csv("data/county/county_general_24.csv")
county_primary_20 <- read_csv("data/county/county_primary_20.csv")
county_primary_20 <- county_primary_20 |>
  rename(County = 1)
county_general_20 <- read_csv("data/county/county_general_20.csv")

#Population Data
county_population <- get_acs(geography="county", 
                       variables = "B01003_001", 
                       state = "MD", year=2022)

#Presidential Election Result (county level)
pres_election_result_county_url <- "https://elections.maryland.gov/elections/2024/general_Results/gen_detail_results_2024_1_1.html"
pres_election_result_county_html <- pres_election_result_county_url |>
  read_html() |>
  html_table()
pres_election_result_county1 <- pres_election_result_county_html[[1]]
pres_election_result_county2 <- pres_election_result_county_html[[2]]
pres_election_result_county2 <- pres_election_result_county2|>
  select(-1)
pres_election_result_county3 <- pres_election_result_county_html[[3]]
pres_election_result_county3 <- pres_election_result_county3|>
  select(-1)
pres_election_result_county4 <- pres_election_result_county_html[[4]]
pres_election_result_county4 <- pres_election_result_county4|>
  select(-1)
pres_election_result_county5 <- pres_election_result_county_html[[5]]
pres_election_result_county5 <- pres_election_result_county5|>
  select(-1)
pres_election_result_county6 <- pres_election_result_county_html[[6]]
pres_election_result_county6 <- pres_election_result_county6|>
  select(-1)
pres_election_result_county <- bind_cols(pres_election_result_county1, pres_election_result_county2, pres_election_result_county3, pres_election_result_county4, pres_election_result_county5, pres_election_result_county6)
```
```{r}
#Primary
#2024
county_primary_24_count <- county_primary_24 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    LIB = sum(LIB),
    OTH = sum(OTH),
    UNA = sum(UNA),
    NLM = sum(NLM)
  ) |>
  mutate(Other = LIB+OTH+NLM) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM24 = 1, REP24 =2, UNA24 = 3, Other24 = 4)
#2020
county_primary_20_count <- county_primary_20 |>
  summarise(
    BAR = sum(BAR),
    DEM = sum(DEM),
    OTH = sum(OTH),
    REP = sum(REP),
    UNA = sum(UNA),
    WCP = sum(WCP)
  ) |>mutate(Other = BAR+OTH+WCP) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM20 = 1, REP20 =2, UNA20 = 3, Other20 = 4)
#Find the change
compare_county_primary_24_20 <- bind_cols(county_primary_24_count, county_primary_20_count) |>
  mutate(DEM_change = ((DEM24-DEM20)/DEM20)*100) |>
  mutate(REP_change = ((REP24-REP20)/REP20)*100) |>
  mutate(UNA_change = ((UNA24-UNA20)/UNA20)*100) |>
  mutate(Other_change = ((Other24-Other20)/Other20)*100)

#General
#2024
county_general_24_count <- county_general_24 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    LIB = sum(LIB),
    OTH = sum(OTH),
    UNA = sum(UNA),
    NLM = sum(NLM),
    GRN = sum(GRN)
  ) |>
  mutate(Other = LIB+OTH+NLM+GRN) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM24 = 1, REP24 =2, UNA24 = 3, Other24 = 4)
#2020
county_general_20_count <- county_general_20 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    BAR = sum(BAR),
    GRN = sum(GRN),
    LIB = sum(LIB),
    WCP = sum(WCP),
    OTH = sum(OTH),
    UNA = sum(UNA)
  ) |>
  mutate(Other = BAR+GRN+LIB+WCP+OTH) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM20 = 1, REP20 =2, UNA20 = 3, Other20 = 4)
#Find the change
compare_county_general_24_20 <- bind_cols(county_general_24_count, county_general_20_count) |>
  mutate(DEM_change = ((DEM24-DEM20)/DEM20)*100) |>
  mutate(REP_change = ((REP24-REP20)/REP20)*100) |>
  mutate(UNA_change = ((UNA24-UNA20)/UNA20)*100) |>
  mutate(Other_change = ((Other24-Other20)/Other20)*100)

#Find Pattern
#Primary24
pattern_county_primary_24 <- county_primary_24 |>
  mutate(Other = LIB+OTH+NLM) |>
  select(County, DEM, REP, UNA, Other, TOTAL) |>
  mutate(DEM_per = (DEM/TOTAL)*100) |>
  mutate(REP_per = (REP/TOTAL)*100) |>
  mutate(UNA_per = (UNA/TOTAL)*100) |>
  mutate(Other_per = (Other/TOTAL)*100) |>
  mutate(DEM_REP = DEM_per-REP_per) |>
  mutate(REP_UNA = REP_per-UNA_per) |>
  mutate(DEM_UNA = DEM_per-UNA_per)
#General24
pattern_county_general_24 <- county_general_24 |>
  mutate(Other = LIB+OTH+NLM+GRN) |>
  select(County, DEM, REP, UNA, Other, TOTAL) |>
  mutate(DEM_per = (DEM/TOTAL)*100) |>
  mutate(REP_per = (REP/TOTAL)*100) |>
  mutate(UNA_per = (UNA/TOTAL)*100) |>
  mutate(Other_per = (Other/TOTAL)*100) |>
  mutate(DEM_REP = DEM_per-REP_per) |>
  mutate(REP_UNA = REP_per-UNA_per) |>
  mutate(DEM_UNA = DEM_per-UNA_per)

#Compare to Population
#Primary24
total_county_primary_24 <- county_primary_24 |>
  select(County, TOTAL) |>
  mutate(County = str_trim(County,side="both"))
cleaned_county_population <- county_population |>
  rename(County = 2, population = 4) |>
  select(2,4) |>
  mutate(County = gsub(" County, Maryland", "", County)) |>
  mutate(County = gsub("Baltimore", "Baltimore County", County)) |>
  mutate(County = gsub("Baltimore County city, Maryland", "Baltimore City", County)) |>
  mutate(County = gsub("St. Mary's", "Saint Mary's", County)) |>
  mutate(County = str_trim(County,side="both"))
regis_population_county_primary_24 <- total_county_primary_24 |>
  left_join(cleaned_county_population, join_by(County)) |>
  mutate(Percent = (TOTAL/population)*100)
#General24
total_county_general_24 <- county_general_24 |>
  select(County, TOTAL) |>
  mutate(County = str_trim(County,side="both"))
regis_population_county_general_24 <- total_county_general_24 |>
  left_join(cleaned_county_population, join_by(County)) |>
  mutate(Percent = (TOTAL/population)*100)

#Compare to the result
#Clean the result
county_result_cleaned <- pres_election_result_county |>
  rename(County = 1, DEM1 = 2, REP1 = 3, LIB = 4, GRN = 5, UNA1 = 6, DEM2 = 7, DEM3 = 8, REP2 = 9, REP3 = 10, OTH1 = 11, OTH2 = 12, OTH3 = 13, OTH4 = 14, OTH5 = 15, OTH6 = 16, UNA2 = 17, UNA3 = 18, UNA4 = 19, UNA5 = 20, UNA6 = 21, UNA7 = 22, UNA8 = 23, UNA9 = 24, UNA10 = 25, UNA11 = 26, UNA12 = 27, UNA13 = 28, UNA14 = 29, UNA15 = 30, UNA16 = 31, UNA17 = 32, UNA18 = 33, UNA19 = 34, OTH7 = 35) |>
  mutate(DEM1 = as.numeric(gsub(",","", DEM1))) |>
  mutate(REP1 = as.numeric(gsub(",","", REP1))) |>
  mutate(LIB = as.numeric(gsub(",","", LIB))) |>
  mutate(GRN = as.numeric(gsub(",","", GRN))) |>
  mutate(UNA1 = as.numeric(gsub(",","", UNA1))) |>
  mutate(DEM2 = as.numeric(gsub(",","", DEM2))) |>
  mutate(DEM3 = as.numeric(gsub(",","", DEM3))) |>
  mutate(REP2 = as.numeric(gsub(",","", REP2))) |>
  mutate(REP3 = as.numeric(gsub(",","", REP3))) |>
  mutate(OTH1 = as.numeric(gsub(",","", OTH1))) |>
  mutate(OTH2 = as.numeric(gsub(",","", OTH2))) |>
  mutate(OTH3 = as.numeric(gsub(",","", OTH3))) |>
  mutate(OTH4 = as.numeric(gsub(",","", OTH4))) |>
  mutate(OTH5 = as.numeric(gsub(",","", OTH5))) |>
  mutate(OTH6 = as.numeric(gsub(",","", OTH6))) |>
  mutate(UNA2 = as.numeric(gsub(",","", UNA2))) |>
  mutate(UNA3 = as.numeric(gsub(",","", UNA3))) |>
  mutate(UNA4 = as.numeric(gsub(",","", UNA4))) |>
  mutate(UNA5 = as.numeric(gsub(",","", UNA5))) |>
  mutate(UNA6 = as.numeric(gsub(",","", UNA6))) |>
  mutate(UNA7 = as.numeric(gsub(",","", UNA7))) |>
  mutate(UNA8 = as.numeric(gsub(",","", UNA8))) |>
  mutate(UNA9 = as.numeric(gsub(",","", UNA9))) |>
  mutate(UNA10 = as.numeric(gsub(",","", UNA10))) |>
  mutate(UNA11 = as.numeric(gsub(",","", UNA11))) |>
  mutate(UNA12 = as.numeric(gsub(",","", UNA12))) |>
  mutate(UNA13 = as.numeric(gsub(",","", UNA13))) |>
  mutate(UNA14 = as.numeric(gsub(",","", UNA14))) |>
  mutate(UNA15 = as.numeric(gsub(",","", UNA15))) |>
  mutate(UNA16 = as.numeric(gsub(",","", UNA16))) |>
  mutate(UNA17 = as.numeric(gsub(",","", UNA17))) |>
  mutate(UNA18 = as.numeric(gsub(",","", UNA18))) |>
  mutate(UNA19 = as.numeric(gsub(",","", UNA19))) |>
  mutate(OTH7 = as.numeric(gsub(",","", OTH7))) |>
  slice(-25) |>
  mutate(DEM = DEM1 + DEM2 + DEM3) |>
  mutate(REP = REP1 + REP2 + REP3) |>
  mutate(UNA = UNA1 + UNA2 + UNA3 + UNA4 + UNA5 + UNA6 + UNA7 + UNA8 + UNA9 + UNA10 + UNA11 + UNA12 + UNA13 + UNA14 + UNA15 + UNA16 + UNA17 + UNA18 + UNA19) |>
  mutate(OTH = OTH1 + OTH2 + OTH3 + OTH4 + OTH5 + OTH6 + OTH7 + LIB + GRN) |>
  select(County, DEM, REP, UNA, OTH) |>
  mutate(Total = DEM+REP+UNA+OTH) |>
  mutate(DEM_per = (DEM/Total)*100) |>
  mutate(REP_per = (REP/Total)*100) |>
  mutate(UNA_per = (UNA/Total)*100) |>
  mutate(OTH_per = (OTH/Total)*100) |>
  mutate(DEM_REP = DEM_per-REP_per) |>
  mutate(REP_UNA = REP_per-UNA_per) |>
  mutate(DEM_UNA = DEM_per-UNA_per)
```

- Has narrative for your code that explains each of your questions, your findings so far and what you do and don't know as a result of them.

*Q1 At the 2024 general election party registration level, several comparisons reveal key voter trends. Comparing registration data from the general election to the primary highlights shifts in  party dynamics, while analyzing changes since 2020 reveal broader trends over time. Relating registrations to the state’s population shows voter participation rates, and comparing registration data to presidential election results provides insights into turnout and party loyalty. Together, these analyses offer a comprehensive view of voter behavior and political engagement in Maryland.*

*Q2 At the congressional district level, we can indicate primary and general voter registrations for each party and compare them to registration data from the previous election in 2020. We can find the pattern of voter registrations in most districts, compare them to the presidential election results pattern, and find the proportion of people who registered to vote compared to all populations in each district.*

*Q3 What we do know*
Closest_races shows the difference between Democratic result and Republican result. The narrow margin indicates a competitive race. We think the close races could be newsworthy because
1. they may lead to recounts or more intense campaigning in future elections
2. the significance of the district may impact the outcome of state elections
3. they may signal a potential shift in voter preferences.
Some districts identified are:
district 33, 30 covering parts of Anne Arundel County
District 38 which covers Somerset County, Worcester County and part of Wicomico County
District 29 covering parts of Calvert County and St. Mary's County
District 2 part of Washington County
District 12, 42 - parts of Baltimore County and Howard County.
District 34 - parts of Harford County.

*What we don't know:*
Historical trends are not part of the current dataset, so we don't know if these close races represent an ongoing trend or a one time occurrence.
We are not sure whether the Unaffiliated (UNF) voters in these districts are playing a decisive role or not. A quick skim led us to believe it is a trivial number but it'd be good to have definite stats applied to it because we know from class that unaffiliated voters are the highest growing party in MD. So far it has not gone beyond 1.4% of the makeup of the voter turnout. The highest is in Talbot County.
We don't have data for all 47 districts in the MD. Our data has granular data on subdivisions within certain districts but not all districts as a whole.*

*Q4 At the precinct level, we cannot work on questions asking about registrations and population due to the lack of data. However, we can find the pattern of presidential election results in each precinct.*

*Q5 Similar to congressional and legislative level, at county level, we can indicate primary and general voter registrations for each party and compare them to registration data from the previous election in 2020. We can find the pattern of voter registrations in most counties, compare them to the presidential election results pattern, and find the proportion of people who registered to vote compared to all populations in each counties. * 


- Identifies your most newsworthy finding so far and explain why it is.

*Q1 Interestingly, while the comparison between 2024 registrations and those of the previous election reveals a significant rise in popularity for "other parties" and "unaffiliated" registrants, a different trend emerges in the context of the 2024 presidential election. The number of actual voters for most parties, including the unaffiliated group, dropped sharply. This suggests that while registrations for secondary parties and unaffiliated voters have grown, when faced with critical decisions in presidential elections, voters tend to gravitate back to the two dominant parties, Republican and Democratic. This pattern highlights a persistent reliance on the traditional bipartisan framework, even as alternative affiliations gain traction in registration data.*

*Q2 At the congressional district level, the voter registration pattern in District 1 could be most newsworthy as it is the only district where people registered for Republican more than Democrats.*

*Q3 District 12, 42 - parts of Baltimore County and Howard County. This was interesting to see because throughout the course of this semester we have seen that these counties lean heavily democrat. It'd be interesting to talk to the residents from these counties and campaign to them.*

*Q4 At the precinct level, in Allegany where Republicans had a big win, precinct ID 2600 could be newsworthy as it was the only area in which Democrats won; similarly, in Montgomery County where Democrats had a big win, precinct ID 107 and 1201 could be newsworthy as they were only two areas in which that Republicans won.*

*Q5 We think the most newsworthy aspect of registration data at the county level is registrations for Republicans and Unaffiliated. Most counties had the same pattern reflecting that registrations for Republicans were more than those for Unaffiliated except in Montgomery, Prince George's, Charles, Frederick, and Howard.*

-----------------------------------------------------------------------------------------------------

*Deliverable 5: The Final Memo and All the Coding Work*

DATA ANALYSIS PROJECT - FINAL MEMO (https://docs.google.com/document/d/1c7djDTlHX2vRUEDavdTbzA_S5vJkOzYXN2ceLLZySo8/edit?usp=sharing)

After the 2024 general election, we'll see many voting trends and patterns related to elections reoccur with numerous races, areas, and many levels of administrative districts in Maryland that confirm our understanding of the area. However, if we look closer, we'll find some variations or outliers among counties, congressional districts, legislative districts, or precincts that contradict what happened in the areas nearby or the general comprehension of the people's behaviors in those areas. These are what we think could lead to interesting stories.

---

*(Story from Q1)*
Unaffiliated Voters Rise in Maryland Registrations but Disappear on Election Day for Presidential Race but not for Senate.

New Findings:

When comparing voter registration to actual voter turnout in Frederick and Calvert counties, stark trends emerge that highlight the dynamics of voter behavior. At the presidential level, "Other Candidates" and unaffiliated voters saw dramatic decreases in actual votes compared to their registration numbers. In Frederick County, unaffiliated voters cast only 5% of the votes relative to their registration, a decline of 95%, while Calvert County exhibited a similar pattern with a 94.75% drop. These findings emphasize that, despite growth in registrations for smaller parties and unaffiliated voters, a significant majority gravitate toward major parties—Republican or Democratic—when making decisive choices in presidential elections. Similarly, smaller parties like the Libertarian and Green parties faced steep declines, with a 27.21% drop in Frederick County and a striking 56.72% in Calvert County.

Interestingly, this trend does not hold at the Senate level. In both counties, smaller parties outperformed expectations based on their registrations. In Frederick County, secondary party votes exceeded registration numbers by 33.07%, and in Calvert County, this increase was even more pronounced at 38.19%. This sharp contrast highlights how partisanship dominates at the presidential level, whereas voters may lean toward alternative options in less nationally polarized contests, such as Senate races.

These results not only confirm the strength of bipartisan loyalty during critical elections but also hint at nuanced voter behavior in less prominent races. This pattern aligns with broader findings from Maryland, where registrations for "Other Candidates" and unaffiliated voters have grown but fail to fully translate into votes during presidential elections.

Final Narrative:

Maryland's elections data reveals a fascinating disconnect between voter registrations and actual turnout, with smaller parties and unaffiliated voters seeing dramatic declines in participation at the presidential level. While smaller parties and unaffiliated groups gained popularity in registration numbers compared to prior elections, their influence sharply decrease when voters faced critical choices.

Maryland's voter registration and turnout data from the 2024 elections reveal compelling insights into voter behavior and the dynamics of party affiliation. Comparing the general election registration totals to the primary elections, both major parties - Democratic and Republican - experienced declines in registrations, while other parties and unaffiliated registrants showed significant increases. These trends also extend to changes since the 2020 elections, with the Democratic Party losing 1.24% in the primary and 2.27% in the general, while unaffiliated registrants rose by 40.24% in the primary and 18.36% in the general. Despite these shifts, overall voter turnout reached 68.23% in the general elections, highlighting strong political engagement relative to the state's population.

The comparison between voter registration and actual turnout in the 2024 presidential election underscores a persistent reliance on major parties during critical decision-making moments. Notably, the Republican Party gained over 19,000 more votes than registered, while the Democratic Party lost over 350,000 voters. Secondary parties and unaffiliated registrants saw sharp drops in actual votes, emphasizing a voter tendency to revert to traditional bipartisan dynamics during national elections, despite the growing popularity of alternative affiliations.

When focusing on Frederick and Calvert counties, the contrast between voter behavior in presidential and Senate races becomes apparent. At the presidential level, unaffiliated voters saw dramatic declines in turnout compared to their registrations, with drops of 95% in Frederick and 94.75% in Calvert. Smaller parties, such as the Libertarian and Green parties, also faced significant losses. However, at the Senate level, secondary parties outperformed expectations, gaining 33.07% more votes than registrations in Frederick and 38.19% in Calvert. These findings suggest that while presidential elections amplify partisanship, Senate races provide opportunities for smaller parties to attract broader support.

In conclusion, Maryland's 2024 election data illustrate a complex interplay between voter registration trends, turnout patterns, and the impact of partisanship across different types of races. While major parties dominate presidential elections, nonpartisan and local contests show more nuanced voter behavior.

--- 

*(Story from Q2)*
Democratic outliers in the Congressional District where Republicans dominate.

In Maryland's Congressional District 1 where most of the people registered to vote for Republicans and it was the only area in Maryland where Republicans won seats in the House, there were variations when looking into each county in this area. Although it seemed like Republicans had a successful performance in District 1 as it had a big win in the Representative race, and Presidential race from the past 2024 general election, there were 3 counties in this area which were Dorchester, Wicomico, and Kent where people registered for Democrats more than Republicans in both 2024 and 2020 elections. We considered taking a closer look at why these 3 counties were the outliers in terms of the characteristics of their population leading to different decision-making in registration to vote could be interesting and newsworthy.

According to the analysis, we got the registration data for each of Maryland's counties from 2024 and 2020. We scoped down to see only counties located in Congressional District 1. We found that every county had registration for Republicans more than Democrats except Dorchester, Wicomico, and Kent which had people registered for Democrats more than Republicans for 0.40%, 2.18%, and 2.55% in 2024 and 5.04%, 5.52%, and 7.50% in 2020 respectively. As we researched, no news story has covered this angle so far.

However, we didn't include Baltimore County in this analysis because some parts of this county were located in Congressional District 2 and we didn't have the exact registration data for only parts of Baltimore County that were in District 1. However, the strength of these data sets was that we could trace the registration data back to 2012 to see if there were trends or changes in registration data in District 1 counties.

To produce this story, we might begin by looking at the registration data in Dorchester, Wicomico, and Kent Counties in the past to confirm if people were always registered for Democrats more than Republicans or if there were any significant changes throughout many years. Then, we could look at https://censusreporter.org/ to get basic information about these 3 counties regarding population, age, race, economic status, etc., and compare them with other counties in District 1. Finally, we could contact Blane H. Miller III, a Democratic candidate who ran for the Representative race in District 1, or his team to ask about the characteristics of voters who voted for him from Dorchester, Wicomico, and Kent Counties which seemed like the majority of votes that he received from the past election.

---

*(Story from Q3)*
Diverging Turnout in Maryland’s District 35

Legislative district-level voter turnout data revealed a very interesting finding. We looked at the change in voter turnout from the general election of 2020 and the general election of 2024. The top four legislative districts with a positive voter turnout change, i.e. more voters in 2024 than 2020 were Legislative District Code 35A, 40, 41, and 45. 

The top four legislative districts with a negative voter turnout change, i.e., more voters in 2020 than in 2024 were, Legislative District Code 09A, 32, 35B, and 42B.

What is interesting here is that Legislative District 35 shows up in both the highest and lowest voter turnout data. District 35 spans Cecil and Harford Counties with the same overall geographic area but with notable turnout disparities.

District 35A:
Vote change: +38,428 votes
Percentage increase: 129%
Population (2020): 45,011
Voting-age population: 35,002
Registered voters: 29,615

District 35B:
Vote change: -31,698 votes
Percentage decrease: -49%
Population (2020): 87,260
Voting-age population: 67,801
Registered voters: 64,272

(Population, voting age population, and registered voters data are taken from Wikipedia. 35A and 35B)
The larger population in 35B makes the 49% vote decrease more striking. 

This could be due to a variety of factors. We are interested in investigating whether it is a data quality issue. Otherwise, 35A and 35B might have different partisan leanings, and voter turnout could vary depending on the candidates or political climate in each sub-district. It could also be due to polling place accessibility. One sub-district could have experienced a surge in voter engagement due to local issues. 

To produce a story, we would be interested in exploring local election campaigns, specific local issues in each sub-district, and population mobility. We could compare these trends to broader regional and state-wide patterns. We are also interested in figuring out whether there were notable redistricting efforts.

---

*(Story from Q4)*
Precincts outliers in counties where Republicans and Democrats had a big win.

In Allegany County where Republicans won the presidential race in most of the precincts, there was 1 precinct, ID "2600", where the Democrats won. This precinct covered the area of the city named "Frostburg". On the contrary, in Montgomery County where Democrats won the same race in most precincts, there were 2 precincts, ID "107" and "1201", where Republicans won. Precincts 107 and 1201 covered the north-western and northern parts of the zip code 20882 area which we understood was a part of the city named "Gaithersburg". We viewed that as these 3 precincts were the outliers of their surrounding areas, telling the story about how they were different from their neighbors in terms of demographic, economic status, or other aspects could be an interesting story.

In our analysis, we looked into presidential election results at the precinct level and indicated which party won in each precinct. Then, we grouped precincts located in the same counties together to see how many precincts the Democrats won and how many the Republicans won in each county. We found that out of 30 precincts in Allegany, "2600" was the only precinct where people voted for Democrats more than Republicans, and out of 257 precincts in Montgomery, "107" and "1201" were only 2 precincts where people voted for Republican more than Democrat.

As we didn't have registration data and population data of these 3 precincts, we might not have been able to compare the election results with registration data to see if they were in line with each other or compare with the population data to see what were the percentage of people who voted for Democrats and who voted for Republicans compared to the whole population, but we could still investigate other aspects.

Nonetheless, before proceeding with the investigation, we should make sure that the territories of these 3 precincts as we understood were correct. We could contact election officials in Allegany and Montgomery and ask them to confirm this information. When we could identify the accurate areas, we could look at https://censusreporter.org/ to get a grasp on basic information about these precincts before conducting the interviews with sources that could answer about demography, economics, or general aspects of these areas and how they were different from nearby neighborhood which could lead to the different decision in presidential voting. We could ask the mayor of the town that these precincts were based in or people who live in these areas.

---

*(Story from Q5)*
Trump's performance in 3 presidential races repeated the same pattern in every Maryland county, except in 6 counties. 

President Donald Trump performed better in the presidential race in the 2024 general election in almost every Maryland county with a higher percentage of votes than those in his previous race in 2020. Nonetheless, this result did not surpass his performance in the 2016 presidential election. This pattern occurred in every county except Baltimore City, Dorchester, Montgomery, Garrett, Prince George's, and Somerset which showed some variations that we considered could lead to interesting stories. 

According to our analysis, we calculated the percentage of votes Trump got from the 2024, 2020, and 2016 races compared to the total votes that he and another main candidate from Democrat in the same race got. We found that in almost every Maryland county, he had a better performance this year compared to what he did in 2020, but it was not on par with what happened in 2016 except in these 6 counties. 

In Baltimore City, Dorchester, and Montgomery, Trump's political comeback this year outperformed what he did in both 2020 and 2016. He gained 12.54% of votes this year compared to 10.91% in 2020 and 11.06% in 2016 for Baltimore City, gained 57.45% of votes this year compared to 56.10% in 2020 and 57.40% in 2016 for Dorchester, and gained 22.56% of votes this year compared to 19.44% in 2020 and 20.58% in 2016 for Montgomery. These are quite interesting, especially in counties that Democrats dominated in many races like Baltimore City and Montgomery.

In Prince George's and Somerset, Trump's performance kept getting better over 3 presidential election races. He gained 8.71% of votes in 2016, 8.91% in 2020, and 11.48% in 2024 for Prince George's which usually is the county that leans to Democrats and gained 56.00% of votes in 2016, 57.51% in 2020, and 58.88% in 2024 for Somerset.

And in Garrett where most of the people voted for Republicans, Trump's performance was declining. He gained 80.76% of votes in 2016, 78.53% in 2020, and down to 77.62% in 2024 while obtained votes in other counties rose.

As we researched, there were some news stories (https://cnsmaryland.org/2024/11/13/trump-makes-inroads-in-true-blue-maryland/) that had already covered this stat, however, they just talked about the result in an overview picture but didn't compare this year's result to the 2016's outcome. Most importantly, we didn't find any stories that talked about these 6 outlier counties specifically. Hence, we thought we could delve a little deeper into them. To produce the stories, we might start by looking at demographic changes in these counties that might explain why voting trends for Trump diverged from other counties. Furthermore, if possible, we might talk to county executives in these areas to get more insightful info about the population. 

---

ALL THE CODING WORK

```{r}
#Load Libraries 
library(tidyverse)
library(janitor)
library(tidycensus)
library(rvest)
library(dplyr)
```

Q1. (new focus) *Bipartisan issue in Presidential and Senate race*

```{r}
#Prep dataframe for Q1 focusing on Maryland
#Registration Data
md_primary_24 <- read_csv("data/county/county_primary_24.csv")
md_general_24 <- read_csv("data/county/county_general_24.csv")
md_primary_20 <- read_csv("data/county/county_primary_20.csv")
md_primary_20 <- md_primary_20 |>
  rename(County = 1)
md_general_20 <- read_csv("data/county/county_general_20.csv")
#Population Data
state_population <- get_acs(geography="state", 
                       variables = "B01003_001", 
                       year=2022)
md_population <- state_population |>
  slice(21)
#Presidential Election Result
pres_election_result_url <- "https://elections.maryland.gov/elections/2024/general_Results/gen_results_2024_1.html"
pres_election_result_html <- pres_election_result_url |>
  read_html() |>
  html_table()
pres_election_result <- pres_election_result_html[[1]]
pres_election_result <- pres_election_result |> clean_names()
pres_election_result <- pres_election_result |>
    slice(-c(34, 35))
```
```{r}
#voters registered for each party in 2024 primary election
total_md_primary_24 <- md_primary_24 |> 
    mutate(OTH = OTH + NLM + LIB) |>
  summarise(
    DEM_tot = sum(DEM),
    REP_tot = sum(REP),
    OTH_tot = sum(OTH),
    UNA_tot = sum(UNA)
  )
```
```{r}
#voters registered for each party in 2024 general election
total_md_general_24 <- md_general_24 |> 
  mutate(OTH = OTH + GRN + NLM + LIB) |> 
summarise(
    DEM_tot = sum(DEM),
    REP_tot = sum(REP),
    OTH_tot = sum(OTH),
    UNA_tot = sum(UNA)
  )
```
```{r}
#voters registered for each party in 2020 primary election
total_md_primary_20 <- md_primary_20 |>
  mutate(OTH = OTH + BAR + WCP) |>
  summarise(
    DEM_tot = sum(DEM),
    REP_tot = sum(REP),
    OTH_tot = sum(OTH),
    UNA_tot = sum(UNA)
  )
```
```{r}
#voters registered for each party in 2020 general election
total_md_general_20 <- md_general_20 |>
  mutate(OTH = OTH + BAR + WCP + GRN + LIB) |>
  summarise(
    DEM_tot = sum(DEM),
    REP_tot = sum(REP),
    OTH_tot = sum(OTH),
    UNA_tot = sum(UNA)
  )
```
```{r}
#How have these numbers changed compared to those in previous elections?*
total_md_primary_24 <- total_md_primary_24 |> 
  mutate(Year = 2024)
total_md_general_24 <- total_md_general_24 |> 
  mutate(Year = 2024)
total_md_primary_20 <- total_md_primary_20 |> 
  mutate(Year = 2020)
total_md_general_20 <- total_md_general_20 |> 
  mutate(Year = 2020)
comparison_general <- bind_rows(total_md_general_24, total_md_general_20) |>
select(Year, everything())
comparison_primary <- bind_rows(total_md_primary_24, total_md_primary_20) |>
select(Year, everything())
```
```{r}
#percentage difference between 2024 and 2022 per each party in geenral elections
comparison_general <- comparison_general|>
    mutate(
    DEM_pct_diff = (DEM_tot[1] - DEM_tot[2]) / DEM_tot[2] * 100,
    REP_pct_diff = (REP_tot[1] - REP_tot[2]) / REP_tot[2] * 100,
    OTH_pct_diff = (OTH_tot[1] - OTH_tot[2]) / OTH_tot[2] * 100,
    UNA_pct_diff = (UNA_tot[1] - UNA_tot[2]) / UNA_tot[2] * 100
  )
```
```{r}
#percentage difference between 2024 and 2022 per each party in primary elections
comparison_primary <- comparison_primary|>
    mutate(
    DEM_pct_diff = (DEM_tot[1] - DEM_tot[2]) / DEM_tot[2] * 100,
    REP_pct_diff = (REP_tot[1] - REP_tot[2]) / REP_tot[2] * 100,
    OTH_pct_diff = (OTH_tot[1] - OTH_tot[2]) / OTH_tot[2] * 100,
    UNA_pct_diff = (UNA_tot[1] - UNA_tot[2]) / UNA_tot[2] * 100
  )
```
```{r}
#voters registered for general 2024 in proportion to population
proportion_md_general_24 <- md_general_24 |>
  summarise(
    total_voters = sum(TOTAL),
    MD_pop = "6161707",
    pct_voters = (total_voters / 6161707) *100
  )
```
```{r}
#voters registered for primary 2024 in proportion to population
proportion_md_primary_24 <- md_primary_24 |>
  summarise(
    total_voters = sum(TOTAL),
    MD_pop = "6161707",
    pct_voters = (total_voters / 6161707) *100
  )
```
```{r}
#*1.3 Compared to the 2024 presidential election result, does the registration pattern match with the result?*
party_elections_results <- pres_election_result |>
  mutate(
    total = as.numeric(gsub(",", "", total))  
  ) |>
  group_by(party) |> 
  summarise(
    Total_votes = sum(total)  
  )
party_elections_results <- party_elections_results |> 
  mutate(
    Total_votes = if_else(party == "Other Candidates", 
                          Total_votes + sum(Total_votes[party %in% c("Green", "Libertarian")]), 
                          Total_votes)
  ) |> 
  filter(!party %in% c("Green", "Libertarian"))
```
```{r}
total_md_general_24_transformed <- total_md_general_24 |> 
  pivot_longer(
    cols = c(DEM_tot, REP_tot, OTH_tot, UNA_tot),  
    names_to = "party",                          
    values_to = "Total_votes"                   
  ) |> 
  mutate(
    party = case_when(                           
      party == "DEM_tot" ~ "Democratic",
      party == "REP_tot" ~ "Republican",
      party == "OTH_tot" ~ "Other Candidates",
      party == "UNA_tot" ~ "Unaffiliated",
      TRUE ~ party
    )
  ) |> 
  select(party, Total_votes)
```
```{r}
comparison_general_elections <- full_join(party_elections_results, total_md_general_24_transformed, by = "party") |>
  rename(Total_election = Total_votes.x, Total_general = Total_votes.y) |>
    mutate(
    Difference = Total_election - Total_general 
    )
```
```{r}
#extract registration results in Frederick

frederick_registration <- md_general_24 |>
  filter(County == "Frederick") |>
  mutate(
    OTH = OTH + LIB + NLM + GRN
  ) |>
  select(-LIB, -NLM, -GRN)
```
```{r}
#extract registration results in Calvert
calvert_registration <- md_general_24 |>
  filter(County == "Calvert") |>
    mutate(
    OTH = OTH + LIB + NLM + GRN
  ) |>
  select(-LIB, -NLM, -GRN)
```
```{r}
# presidential election Frederick data
frederick_pres_results <- tibble::tibble(
  County = "Frederick",
  DEM = 82409,
  REP = 68753,
  UNA = 1494 + 1110,
  OTH = 1378 + 970,
  TOTAL = 82409 + 68753 + 970 + 1378 + 1494 + 1110
)
```
```{r}
#presidential election Calvert data

calvert_pres_results <- read.csv("data/county/calvert_presidential_off.csv", header = FALSE) 
colnames(calvert_pres_results) <- c("candidate", "party", "Early Voting",	"Election Day",	"Mail-In", "Ballot	Provisional", "Total", "Percentage")
```
```{r}
#Calvert presidential

calvert_pres_results <- calvert_pres_results|>
mutate(across(everything(), ~ replace_na(., "Unaffiliated")))


#clean Calvert presidential
calvert_pres_results_clean <- calvert_pres_results |>
  mutate(
    Total = as.numeric(gsub(",", "", Total)), 
  ) |>
  group_by(party) |> 
  summarise(
    Total_votes = sum(Total, na.rm = TRUE)  
  )|>
  slice(-1)

#Add Write-ins vote to Unaffiliated - Calvert presidential
calvert_pres_results_clean <- calvert_pres_results_clean |> 
   mutate(
       Total_votes = if_else(party == "Unaffiliated", Total_votes + 274, Total_votes),
  )

#Summing the votes of Green, Libertarian - Calvert presidential
calvert_pres_results_final <- calvert_pres_results_clean |>
  bind_rows(
    calvert_pres_results_clean |> 
      filter(party %in% c("Green", "Libertarian")) |> 
      summarise(
        party = "Other candidates", 
        Total_votes = sum(Total_votes, na.rm = TRUE)  # Sum of votes for these parties
      )
  )

#Adding Other Candidates to Other candidate - Calvert presidential
calvert_pres_results_final <- calvert_pres_results_final |> 
   mutate(
       Total_votes = if_else(party == "Other candidates", Total_votes + 8, Total_votes),
  )|>
  slice(-2,-3,-4) |>
  mutate(party = if_else(party == "Other candidates", "Other Candidates", party))

```
```{r}
#senate election Frederick data

frederick_senate_results <- tibble::tibble(
  County = "Frederick",
  DEM =  66280,
  REP = 83767,
  UNA = 2 + 348 + 79,
  OTH = 4293,
  TOTAL =  66280 + 83767 + 4293 + 79 + 2 + 348 
)
```
```{r}
#senate election Calvert data

calvert_senate_results <- read.csv("data/county/calvert_senate_off.csv", header = FALSE)
colnames(calvert_senate_results) <- c("candidate", "party", "Early Voting",	"Election Day",	"Mail-In", "Ballot	Provisional", "Total", "Percentage")

```
```{r}
#Calvert senate

calvert_senate_results <- calvert_senate_results|>
mutate(across(everything(), ~ replace_na(., "Unaffiliated")))

#clean Calvert Senate

calvert_senate_results_clean <- calvert_senate_results |>
  mutate(
    Total = as.numeric(gsub(",", "", Total)), 
  ) |>
  group_by(party) |> 
  summarise(
    Total_votes = sum(Total, na.rm = TRUE)  
  )|>
  slice(-1)

#Add Write-ins vote to Unaffiliated - Calvert senate
calvert_senate_results_clean <- calvert_senate_results_clean |> 
   mutate(
       Total_votes = if_else(party == "Unaffiliated", Total_votes + 95, Total_votes),
  )

#Summing the votes of Libertarian and Green to other candidates - Calvert senate
calvert_senate_results_final <- calvert_senate_results_clean |>
  bind_rows(
    calvert_senate_results_clean |> 
      filter(party %in% c("Green","Libertarian")) |> 
      summarise(
        party = "Other candidates", 
        Total_votes = sum(Total_votes, na.rm = TRUE)  # Sum of votes for these parties
      )
  )

#Adding Other Candidates to Other candidate - Calvert senate
calvert_senate_results_final <- calvert_senate_results_final |> 
   mutate(
       Total_votes = if_else(party == "Other candidates", Total_votes + 24, Total_votes),
  )|>
  slice(-2,-3) |>
  mutate(party = if_else(party == "Other candidates", "Other Candidates", party))
```
```{r}
#transform Frederick registration results

frederick_registration_trans <- frederick_registration |> 
  pivot_longer(
    cols = c(DEM, REP, OTH, UNA),  
    names_to = "party",                          
    values_to = "Total_votes"                   
  ) |> 
  mutate(
    party = case_when(
      party == "DEM" ~ "Democratic",
      party == "REP" ~ "Republican",
      party == "UNA" ~ "Unaffiliated",
      party == "OTH" ~ "Other Candidates",
      TRUE ~ party
    )
  ) |>
  select(party, Total_votes)
```
```{r}
#transform Calvert registration results
calvert_registration_trans <- calvert_registration |> 
  pivot_longer(
    cols = c(DEM, REP, OTH, UNA),  
    names_to = "party",                          
    values_to = "Total_votes"                   
  ) |> 
  mutate(
    party = case_when(
      party == "DEM" ~ "Democratic",
      party == "REP" ~ "Republican",
      party == "UNA" ~ "Unaffiliated",
      party == "OTH" ~ "Other Candidates",
      TRUE ~ party
    )
  ) |>
  select(party, Total_votes)

```
```{r}
#transform Frederick presidential results
frederick_pres_results_trans <- frederick_pres_results |>
  pivot_longer(
    cols = c(DEM, REP, OTH, UNA),  
    names_to = "party",                          
    values_to = "Total_votes"                   
  ) |> 
  mutate(
    party = case_when(                           
      party == "DEM" ~ "Democratic",
      party == "REP" ~ "Republican",
      party == "OTH" ~ "Other Candidates",
      party == "UNA" ~ "Unaffiliated",
      TRUE ~ party
    )
  ) |> 
  select(party, Total_votes)
```
```{r}
#transform Frederick senate results

frederick_senate_trans <- frederick_senate_results |> 
  pivot_longer(
    cols = c(DEM, REP, OTH, UNA),  
    names_to = "party",                          
    values_to = "Total_votes"                   
  ) |> 
  mutate(
    party = case_when(
      party == "DEM" ~ "Democratic",
      party == "REP" ~ "Republican",
      party == "UNA" ~ "Unaffiliated",
      party == "OTH" ~ "Other Candidates",
      TRUE ~ party
    )
  ) |>
  select(party, Total_votes)
```
```{r}
#comparison presidential and registrations Frederick

comparison_frederick_presidential <- full_join(frederick_pres_results_trans, frederick_registration_trans, by = "party") |>
  rename(Total_presidential = Total_votes.x, Total_registration = Total_votes.y) |>
    mutate(
    Difference = Total_presidential - Total_registration,
    Percent_difference = (Difference / Total_registration) * 100
    )


#comparison senate and registration Frederick

comparison_frederick_senate <- full_join(frederick_senate_trans, frederick_registration_trans, by = "party") |>
  rename(Total_senate = Total_votes.x, Total_registration = Total_votes.y) |>
    mutate(
    Difference = Total_senate - Total_registration,
    Percent_difference = (Difference / Total_registration) * 100
    )


#comparison presidential and registrations Calvert

comparison_calvert_presidential <- full_join(calvert_pres_results_final, calvert_registration_trans, by = "party") |>
  rename(Total_presidential = Total_votes.x, Total_registration = Total_votes.y) |>
    mutate(
    Difference = Total_presidential - Total_registration,
    Percent_difference = (Difference / Total_registration) * 100
    )

#comparison senate and registration Calvert

comparison_calvert_senate <- full_join(calvert_senate_results_final, calvert_registration_trans, by = "party") |>
  rename(Total_senate = Total_votes.x, Total_registration = Total_votes.y) |>
    mutate(
    Difference = Total_senate - Total_registration,
    Percent_difference = (Difference / Total_registration) * 100
    )

```

Q2. *Congressional District level*

```{r}
#Prep dataframe for Q2 focusing on congressional district level

#Registration Data
congress_primary_24 <- read_csv("data/congress_level/congress_primary_24.csv")
congress_general_24 <- read_csv("data/congress_level/congress_general_24.csv")
congress_primary_20 <- read_csv("data/congress_level/congress_primary_20.csv")
congress_primary_20 <- congress_primary_20 |>
  rename(District = 1) |>
  rename(county = 2) |>
  rename(BAR = 3) |>
  rename(DEM = 4) |>
  rename(OTH = 5) |>
  rename(REP = 6) |>
  rename(UNA = 7) |>
  rename(WCP = 8) |>
  rename(total = 9) 
congress_general_20 <- read_csv("data/congress_level/congress_general_20.csv")

#Population Data
congress_population <- get_acs(geography="congressional district", 
                       variables = "B01003_001", state = "MD", 
                       year=2022)

#Representative Election Result
representative_election_result_url <- "https://elections.maryland.gov/elections/2024/general_Results/gen_results_2024_3.html"
representative_election_result_html <- representative_election_result_url |>
  read_html() |>
  html_table()
representative_election_result1 <- representative_election_result_html[[1]]
representative_election_result2 <- representative_election_result_html[[2]]
representative_election_result3 <- representative_election_result_html[[3]]
representative_election_result4 <- representative_election_result_html[[4]]
representative_election_result5 <- representative_election_result_html[[5]]
representative_election_result6 <- representative_election_result_html[[6]]
representative_election_result7 <- representative_election_result_html[[7]]
representative_election_result8 <- representative_election_result_html[[8]]
```
```{r}
#Primary
#2024
count_congress_primary_24 <- congress_primary_24 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    LIB = sum(LIB),
    OTH = sum(OTH),
    UNA = sum(UNA),
    NLM = sum(NLM)
  ) |>
  mutate(Other = LIB+OTH+NLM) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM24 = 1, REP24 =2, UNA24 = 3, Other24 = 4)
#2020
count_congress_primary_20 <- congress_primary_20 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    OTH = sum(OTH),
    UNA = sum(UNA),
    BAR = sum(BAR),
    WCP = sum(WCP)
  ) |>
  mutate(Other = OTH+BAR+WCP) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM20 = 1, REP20 =2, UNA20 = 3, Other20 = 4)
#Find the change
compare_congress_primary_24_20 <- bind_cols(count_congress_primary_24, count_congress_primary_20) |>
  mutate(DEM_change = ((DEM24-DEM20)/DEM20)*100) |>
  mutate(REP_change = ((REP24-REP20)/REP20)*100) |>
  mutate(UNA_change = ((UNA24-UNA20)/UNA20)*100) |>
  mutate(Other_change = ((Other24-Other20)/Other20)*100)

#General
#2024
count_congress_general_24 <- congress_general_24 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    LIB = sum(LIB),
    OTH = sum(OTH),
    UNA = sum(UNA),
    NLM = sum(NLM),
    GRN = sum(GRN)
  ) |>
  mutate(Other = LIB+OTH+NLM+GRN) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM24 = 1, REP24 =2, UNA24 = 3, Other24 = 4)
#2020
count_congress_general_20 <- congress_general_20 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    BAR = sum(BAR),
    GRN = sum(GRN),
    LIB = sum(LIB),
    WCP = sum(WCP),
    OTH = sum(OTH),
    UNA = sum(UNA)
  ) |>
  mutate(Other = OTH+BAR+GRN+LIB+WCP) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM20 = 1, REP20 =2, UNA20 = 3, Other20 = 4)
#Find the change
compare_congress_general_24_20 <- bind_cols(count_congress_general_24, count_congress_general_20) |>
  mutate(DEM_change = ((DEM24-DEM20)/DEM20)*100) |>
  mutate(REP_change = ((REP24-REP20)/REP20)*100) |>
  mutate(UNA_change = ((UNA24-UNA20)/UNA20)*100) |>
  mutate(Other_change = ((Other24-Other20)/Other20)*100)

#Find Pattern
#Primary24
pattern_congress_primary_24 <- congress_primary_24 |>
  mutate(Other = LIB+OTH+NLM) |>
  select(District, DEM, REP, UNA, Other, TOTAL) |>
  mutate(DEM_per = (DEM/TOTAL)*100) |>
  mutate(REP_per = (REP/TOTAL)*100) |>
  mutate(UNA_per = (UNA/TOTAL)*100) |>
  mutate(Other_per = (Other/TOTAL)*100)
#General24
pattern_congress_general_24 <- congress_general_24 |>
  mutate(Other = LIB+OTH+NLM+GRN) |>
  select(District, DEM, REP, UNA, Other, TOTAL) |>
  mutate(DEM_per = (DEM/TOTAL)*100) |>
  mutate(REP_per = (REP/TOTAL)*100) |>
  mutate(UNA_per = (UNA/TOTAL)*100) |>
  mutate(Other_per = (Other/TOTAL)*100)

#Compare to Population
#Primary24
total_congress_primary_24 <- congress_primary_24 |>
  select(District, TOTAL)
cleaned_congress_population <- congress_population |>
  rename(population = 4) |>
  select(4)
regis_population_congress_primary_24 <- bind_cols(total_congress_primary_24, cleaned_congress_population) |>
  mutate(Percent = (TOTAL/population)*100)
#General24
total_congress_general_24 <- congress_general_24 |>
  select(District, TOTAL)
regis_population_congress_general_24 <- bind_cols(total_congress_general_24, cleaned_congress_population) |>
  mutate(Percent = (TOTAL/population)*100)

#Compare to the result
#Clean the result
representative_election_result1 <- representative_election_result1 |>
  slice(-5) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, x = 4, y = 5) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(x = as.numeric(gsub("%","", x))) |>
  mutate(y = as.numeric(gsub("%","", y))) |>
  mutate(Other_percent_result = x+y) |>
  select(1,2,3,6)
representative_election_result2 <- representative_election_result2 |>
  slice(-5) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, x = 4, y = 5) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(x = as.numeric(gsub("%","", x))) |>
  mutate(y = as.numeric(gsub("%","", y))) |>
  mutate(Other_percent_result = x+y) |>
  select(1,2,3,6)
representative_election_result3 <- representative_election_result3 |>
  slice(-5) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, x = 4, y = 5) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(x = as.numeric(gsub("%","", x))) |>
  mutate(y = as.numeric(gsub("%","", y))) |>
  mutate(Other_percent_result = x+y) |>
  select(1,2,3,6)
representative_election_result4 <- representative_election_result4 |>
  slice(-4) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, Other_percent_result = 4) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(Other_percent_result = as.numeric(gsub("%","", Other_percent_result)))
representative_election_result5 <- representative_election_result5 |>
  slice(-4) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, Other_percent_result = 4) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(Other_percent_result = as.numeric(gsub("%","", Other_percent_result)))
representative_election_result6 <- representative_election_result6 |>
  slice(-4) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, Other_percent_result = 4) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(Other_percent_result = as.numeric(gsub("%","", Other_percent_result)))
representative_election_result7 <- representative_election_result7 |>
  slice(-5) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, x = 4, y = 5) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(x = as.numeric(gsub("%","", x))) |>
  mutate(y = as.numeric(gsub("%","", y))) |>
  mutate(Other_percent_result = x+y) |>
  select(1,2,3,6)
representative_election_result8 <- representative_election_result8 |>
  slice(-5) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, x = 4, y = 5) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(x = as.numeric(gsub("%","", x))) |>
  mutate(y = as.numeric(gsub("%","", y))) |>
  mutate(Other_percent_result = x+y) |>
  select(1,2,3,6)
representative_election_result <- bind_rows(list(representative_election_result1, representative_election_result2, representative_election_result3, representative_election_result4, representative_election_result5, representative_election_result6, representative_election_result7, representative_election_result8)) |>
  select(2,3,4)
#Merge with the registration
congress_general_regis_result <- bind_cols(pattern_congress_general_24, representative_election_result) |>
  select(-2,-3,-4,-5,-6)
```
```{r}
#Looking at counties in Congressional District 1

#See the variations of registration data in District 1's counties  
#Registration Data
county_general_24 <- read_csv("data/county/county_general_24.csv")
county_general_20 <- read_csv("data/county/county_general_20.csv")
#General24
pattern_county_general_24 <- county_general_24 |>
  mutate(Other = LIB+OTH+NLM+GRN) |>
  select(County, DEM, REP, UNA, Other, TOTAL) |>
  mutate(DEM_per = (DEM/TOTAL)*100) |>
  mutate(REP_per = (REP/TOTAL)*100) |>
  mutate(UNA_per = (UNA/TOTAL)*100) |>
  mutate(Other_per = (Other/TOTAL)*100) |>
  mutate(DEM_REP = DEM_per-REP_per) |>
  mutate(REP_UNA = REP_per-UNA_per) |>
  mutate(DEM_UNA = DEM_per-UNA_per)
#Filter only counties in District 1
District1_pattern_county_general_24 <- pattern_county_general_24 |>
  filter(County %in% c("Baltimore County", "Caroline", "Cecil", "Dorchester", "Harford", "Kent", "Queen Anne's", "Somerset", "Talbot", "Wicomico", "Worcester"))
#General20
pattern_county_general_20 <- county_general_20 |>
  mutate(Other = BAR+GRN+LIB+WCP+OTH) |>
  select(County, DEM, REP, UNA, Other, TOTAL) |>
  mutate(DEM_per = (DEM/TOTAL)*100) |>
  mutate(REP_per = (REP/TOTAL)*100) |>
  mutate(UNA_per = (UNA/TOTAL)*100) |>
  mutate(Other_per = (Other/TOTAL)*100) |>
  mutate(DEM_REP = DEM_per-REP_per) |>
  mutate(REP_UNA = REP_per-UNA_per) |>
  mutate(DEM_UNA = DEM_per-UNA_per)
#Filter only counties in District 1
District1_pattern_county_general_20 <- pattern_county_general_20 |>
  filter(County %in% c("Baltimore County", "Caroline", "Cecil", "Dorchester", "Harford", "Kent", "Queen Anne's", "Somerset", "Talbot", "Wicomico", "Worcester"))

#See the results of Representative race and Presidential race in District 1's counties 
#Representative race
#Prep Data
representative1_election_result_url <- "https://elections.maryland.gov/elections/2024/general_Results/gen_detail_results_2024_3_1_District%201.html"
representative1_election_result_html <- representative1_election_result_url |>
  read_html() |>
  html_table()
representative1_election_result1 <- representative1_election_result_html[[1]]
#Clean
District1_representative1_election_result1 <- representative1_election_result1 |>
  slice(-12)|>
  rename(County = 1, DEM = 2, REP = 3, LIB = 4, Write = 5) |>
  mutate(DEM = as.numeric(gsub(",","",DEM))) |>
  mutate(REP = as.numeric(gsub(",","",REP))) |>
  mutate(LIB = as.numeric(gsub(",","",LIB))) |>
  mutate(Write = as.numeric(gsub(",","",Write))) |>
  mutate(Other = LIB + Write) |>
  select(County, DEM, REP, Other) |>
  mutate(Total = DEM + REP + Other) |>
  mutate(DEM_per = (DEM/Total)*100) |>
  mutate(REP_per = (REP/Total)*100) |>
  mutate(DEP_REP = DEM_per - REP_per)

#Presidential Race
#Prep Data
pres_election_result_county_url <- "https://elections.maryland.gov/elections/2024/general_Results/gen_detail_results_2024_1_1.html"
pres_election_result_county_html <- pres_election_result_county_url |>
  read_html() |>
  html_table()
pres_election_result_county1 <- pres_election_result_county_html[[1]]
pres_election_result_county2 <- pres_election_result_county_html[[2]]
pres_election_result_county2 <- pres_election_result_county2|>
  select(-1)
pres_election_result_county3 <- pres_election_result_county_html[[3]]
pres_election_result_county3 <- pres_election_result_county3|>
  select(-1)
pres_election_result_county4 <- pres_election_result_county_html[[4]]
pres_election_result_county4 <- pres_election_result_county4|>
  select(-1)
pres_election_result_county5 <- pres_election_result_county_html[[5]]
pres_election_result_county5 <- pres_election_result_county5|>
  select(-1)
pres_election_result_county6 <- pres_election_result_county_html[[6]]
pres_election_result_county6 <- pres_election_result_county6|>
  select(-1)
pres_election_result_county <- bind_cols(pres_election_result_county1, pres_election_result_county2, pres_election_result_county3, pres_election_result_county4, pres_election_result_county5, pres_election_result_county6)
#Clean
county_result_cleaned <- pres_election_result_county |>
  rename(County = 1, DEM1 = 2, REP1 = 3, LIB = 4, GRN = 5, UNA1 = 6, DEM2 = 7, DEM3 = 8, REP2 = 9, REP3 = 10, OTH1 = 11, OTH2 = 12, OTH3 = 13, OTH4 = 14, OTH5 = 15, OTH6 = 16, UNA2 = 17, UNA3 = 18, UNA4 = 19, UNA5 = 20, UNA6 = 21, UNA7 = 22, UNA8 = 23, UNA9 = 24, UNA10 = 25, UNA11 = 26, UNA12 = 27, UNA13 = 28, UNA14 = 29, UNA15 = 30, UNA16 = 31, UNA17 = 32, UNA18 = 33, UNA19 = 34, OTH7 = 35) |>
  mutate(DEM1 = as.numeric(gsub(",","", DEM1))) |>
  mutate(REP1 = as.numeric(gsub(",","", REP1))) |>
  mutate(LIB = as.numeric(gsub(",","", LIB))) |>
  mutate(GRN = as.numeric(gsub(",","", GRN))) |>
  mutate(UNA1 = as.numeric(gsub(",","", UNA1))) |>
  mutate(DEM2 = as.numeric(gsub(",","", DEM2))) |>
  mutate(DEM3 = as.numeric(gsub(",","", DEM3))) |>
  mutate(REP2 = as.numeric(gsub(",","", REP2))) |>
  mutate(REP3 = as.numeric(gsub(",","", REP3))) |>
  mutate(OTH1 = as.numeric(gsub(",","", OTH1))) |>
  mutate(OTH2 = as.numeric(gsub(",","", OTH2))) |>
  mutate(OTH3 = as.numeric(gsub(",","", OTH3))) |>
  mutate(OTH4 = as.numeric(gsub(",","", OTH4))) |>
  mutate(OTH5 = as.numeric(gsub(",","", OTH5))) |>
  mutate(OTH6 = as.numeric(gsub(",","", OTH6))) |>
  mutate(UNA2 = as.numeric(gsub(",","", UNA2))) |>
  mutate(UNA3 = as.numeric(gsub(",","", UNA3))) |>
  mutate(UNA4 = as.numeric(gsub(",","", UNA4))) |>
  mutate(UNA5 = as.numeric(gsub(",","", UNA5))) |>
  mutate(UNA6 = as.numeric(gsub(",","", UNA6))) |>
  mutate(UNA7 = as.numeric(gsub(",","", UNA7))) |>
  mutate(UNA8 = as.numeric(gsub(",","", UNA8))) |>
  mutate(UNA9 = as.numeric(gsub(",","", UNA9))) |>
  mutate(UNA10 = as.numeric(gsub(",","", UNA10))) |>
  mutate(UNA11 = as.numeric(gsub(",","", UNA11))) |>
  mutate(UNA12 = as.numeric(gsub(",","", UNA12))) |>
  mutate(UNA13 = as.numeric(gsub(",","", UNA13))) |>
  mutate(UNA14 = as.numeric(gsub(",","", UNA14))) |>
  mutate(UNA15 = as.numeric(gsub(",","", UNA15))) |>
  mutate(UNA16 = as.numeric(gsub(",","", UNA16))) |>
  mutate(UNA17 = as.numeric(gsub(",","", UNA17))) |>
  mutate(UNA18 = as.numeric(gsub(",","", UNA18))) |>
  mutate(UNA19 = as.numeric(gsub(",","", UNA19))) |>
  mutate(OTH7 = as.numeric(gsub(",","", OTH7))) |>
  slice(-25) |>
  mutate(DEM = DEM1 + DEM2 + DEM3) |>
  mutate(REP = REP1 + REP2 + REP3) |>
  mutate(UNA = UNA1 + UNA2 + UNA3 + UNA4 + UNA5 + UNA6 + UNA7 + UNA8 + UNA9 + UNA10 + UNA11 + UNA12 + UNA13 + UNA14 + UNA15 + UNA16 + UNA17 + UNA18 + UNA19) |>
  mutate(OTH = OTH1 + OTH2 + OTH3 + OTH4 + OTH5 + OTH6 + OTH7 + LIB + GRN) |>
  select(County, DEM, REP, UNA, OTH) |>
  mutate(Total = DEM+REP+UNA+OTH) |>
  mutate(DEM_per = (DEM/Total)*100) |>
  mutate(REP_per = (REP/Total)*100) |>
  mutate(UNA_per = (UNA/Total)*100) |>
  mutate(OTH_per = (OTH/Total)*100) |>
  mutate(DEM_REP = DEM_per-REP_per) |>
  mutate(REP_UNA = REP_per-UNA_per) |>
  mutate(DEM_UNA = DEM_per-UNA_per)
#Scope down to only counties on District 1
District1_county_result_cleaned <- county_result_cleaned |>
  filter(County %in% c("Baltimore County", "Caroline", "Cecil", "Dorchester", "Harford", "Kent", "Queen Anne's", "Somerset", "Talbot", "Wicomico", "Worcester"))
```

Q3. *Legislative District level*

```{r}
#Prep dataframe for Q3 focusing on legislative district level

#Registration Data
legis_primary_24 <- read_csv("data/legis_level/legis_primary_24.csv")
legis_general_24 <- read_csv("data/legis_level/legis_general_24.csv")
legis_primary_20 <- read_csv("data/legis_level/legis_primary_20.csv")
legis_general_20 <- read_csv("data/legis_level/legis_general_20.csv")

#Population Data
legis_population <- get_acs(geography="state legislative district (lower chamber)", 
                       variables = "B01003_001", state = "MD", 
                       year=2022)

#Presidential Election Result 
#Download Precinct Result and Code
precint_result <- read_csv("data/legis_level/result/result.csv")
legis_code <- read_csv("data/legis_level/result/code.csv")
#Prep result dataframe
precint_result <- precint_result |>
  filter(Contest == "President and Vice President of the United States")
precint_result <- precint_result |>
  select(2,4,6,7) |>
  clean_names() |>
  mutate(precinct_name = as.numeric(gsub("-","", precinct_name)))
#Prep code dataframe
legis_code <- legis_code |>
  select(3,4,10) |>
  filter(!is.na(ED_PRECINCT)) |>
  filter(!is.na(LEGISLATIVE_SUB_DISTRICT_CODE))
#Join result dataframe with code dataframe to generate legislative district results 
legis_result_draft <- precint_result |> 
  left_join(legis_code, join_by(jurisdiction==COUNTY_NAME, precinct_name==ED_PRECINCT)) |>
  filter(!is.na(LEGISLATIVE_SUB_DISTRICT_CODE)) |>
  group_by(LEGISLATIVE_SUB_DISTRICT_CODE, party) |>
  summarise(
    total_votes = sum(total_votes))
#Reshape dataframe
legis_result_DEM <- legis_result_draft |>
  filter(party == "DEM") |>
  select(1,3) |>
  rename(DEM_result = 2)
legis_result_GRN <- legis_result_draft |>
  filter(party == "GRN") |>
  select(3) |>
  rename(GRN_result = 2) 
legis_result_LIB <- legis_result_draft |>
  filter(party == "LIB") |>
  select(3) |>
  rename(LIB_result = 2)
legis_result_NON <- legis_result_draft |>
  filter(party == "NON") |>
  select(3) |>
  rename(NON_result = 2)
legis_result_REP <- legis_result_draft |>
  filter(party == "REP") |>
  select(3) |>
  rename(REP_result = 2)
legis_result_UNF <- legis_result_draft |>
  filter(party == "UNF") |>
  select(3) |>
  rename(UNF_result = 2)
legis_result <- bind_cols(legis_result_DEM, legis_result_GRN, legis_result_LIB, legis_result_NON, legis_result_REP, legis_result_UNF) |>
  select(-3,-5,-7,-9,-11) |>
  rename(legislative_district_code = 1)

```
```{r}
legis_result <- legis_result |>
  mutate(dem_rep_difference = DEM_result - REP_result)|>
  mutate(total_votes = DEM_result + GRN_result + LIB_result + NON_result + REP_result + UNF_result)|>
  arrange(desc(dem_rep_difference))

legis_result <- legis_result |>
  mutate(
    party_dominance = case_when(
      dem_rep_difference > 0 ~ "Democrat Dominant",
      dem_rep_difference < 0 ~ "Republican Dominant",
      TRUE ~ "No Clear Winner" # In case of a tie or no votes
    )
  )
dominance_counts <- legis_result |>
  count(party_dominance)
print(dominance_counts)

closest_races <- legis_result |>
  select(legislative_district_code, DEM_result, REP_result) |>
  mutate(margin = abs(DEM_result - REP_result)) |>
  arrange(margin) |>
  head(10) 

legis_result <- legis_result |>
  mutate(
    unf_percentage = (UNF_result / total_votes) * 100 
  )


legis_result_sorted <- legis_result |>
  arrange(desc(unf_percentage))


head(legis_result_sorted, 10)
```
```{r}
legis_general_20 <- read_csv("data/legis_level/legis_general_20.csv")
legis_general_24 <- read_csv("data/legis_level/legis_general_24.csv")


legis_general_20 <- legis_general_20 |>
  mutate(year = 2020)

legis_general_24 <- legis_general_24 |>
  mutate(year = 2024)


combined_data_kriti <- legis_general_20 |>
  full_join(legis_general_24, by = c("District", "County"), suffix = c("_20", "_24"))

combined_data_kriti <- combined_data_kriti |>
  mutate(
    turnout_change = if_else(
      !is.na(Total_20) & !is.na(Total_24),  #if both columns are non  missing
      Total_24 - Total_20,
      NA_real_  # if either Total_20 or Total_24 is missing leaving as Na
    ),
    percentage_change = if_else(
      Total_20 > 0,  #Avoiding division by zero
      (Total_24 - Total_20)/Total_20 * 100,
      NA_real_  
    )
  )|>
  arrange(desc(turnout_change))

combined_data_wona<- combined_data_kriti |> # without NA values table
  drop_na()

head(combined_data_wona, 4)|>
  select( District,Total_24, Total_20, turnout_change,percentage_change)

tail(combined_data_wona, 4)|>
  select( District,Total_24, Total_20, turnout_change, percentage_change)
```

Q4.*Precinct level*

```{r}
#Prep dataframe for Q4 focusing on legislative precinct level

#Presidential Result Data
precinct_result <- read_csv("data/precinct/precinct result.csv")
precinct_result_DEM <- precinct_result |>
  select(1, 2, 3, 4, 6, 7) |>
  filter(Contest == "President and Vice President of the United States") |>
  select(-1) |>
  filter(Party == "DEM") |>
  rename(total_votes_DEM = 5) |>
  select(-4)
precinct_result_REP <- precinct_result |>
  select(1, 2, 3, 4, 6, 7) |>
  filter(Contest == "President and Vice President of the United States") |>
  select(-1) |>
  filter(Party == "REP") |>
  rename(total_votes_REP = 5) |>
  select(-1,-2,-3,-4)
precinct_result_LIB <- precinct_result |>
  select(1, 2, 3, 4, 6, 7) |>
  filter(Contest == "President and Vice President of the United States") |>
  select(-1) |>
  filter(Party == "LIB") |>
  rename(total_votes_LIB = 5) |>
  select(-1,-2,-3,-4)
precinct_result_GRN <- precinct_result |>
  select(1, 2, 3, 4, 6, 7) |>
  filter(Contest == "President and Vice President of the United States") |>
  select(-1) |>
  filter(Party == "GRN") |>
  rename(total_votes_GRN = 5) |>
  select(-1,-2,-3,-4)
precinct_result_UNF <- precinct_result |>
  select(1, 2, 3, 4, 6, 7) |>
  filter(Contest == "President and Vice President of the United States") |>
  select(-1) |>
  filter(Party == "UNF") |>
  rename(total_votes_UNF = 5) |>
  select(-1,-2,-3,-4)
precinct_result_NON <- precinct_result |>
  select(1, 2, 3, 4, 6, 7) |>
  filter(Contest == "President and Vice President of the United States") |>
  select(-1) |>
  filter(Party == "NON") |>
  rename(total_votes_NON = 5) |>
  select(-1,-2,-3,-4)
precinct_result_cleaned <- bind_cols(precinct_result_DEM, precinct_result_REP, precinct_result_LIB, precinct_result_GRN, precinct_result_UNF, precinct_result_NON)
precinct_result_final <- precinct_result_cleaned |>
  mutate(total_votes = total_votes_DEM + total_votes_REP + total_votes_LIB + total_votes_GRN + total_votes_UNF + total_votes_NON)
```
```{r}
#Find the pattern for the results
precinct_result_pattern <- precinct_result_final |>
  mutate(total_votes_others = total_votes_LIB + total_votes_GRN + total_votes_NON) |>
  select(1,2,3,4,5,11,8,10,) |>
  filter(total_votes != 0) |>
  mutate(DEM_percent = (total_votes_DEM/total_votes)*100) |>
  mutate(REP_percent = (total_votes_REP/total_votes)*100) |>
  mutate(others_percent = (total_votes_others/total_votes)*100) |>
  mutate(UNF_percent = (total_votes_UNF/total_votes)*100) |>
  mutate(DEM_REP = DEM_percent - REP_percent) |>
  select(1,2,3,13) |>
  filter(DEM_REP != 0)

#Check each county
precinct_result_pattern_compare <- precinct_result_pattern |>
  mutate(
    D = case_when(
      DEM_REP > 0 ~ 1,
      DEM_REP < 0 ~ 0
      ),
    R = case_when(
      DEM_REP > 0 ~ 0,
      DEM_REP < 0 ~ 1
      )
  ) |>
  group_by(Jurisdiction)|>
  summarise(
    county_count = n(),
    More_DEM = sum(D),
    More_REP = sum(R)
  ) |>
  mutate(More_DEM_percent = (More_DEM/county_count)*100) |>
  mutate(More_REP_percent = (More_REP/county_count)*100)
```

Q5. *County Level* 

```{r}
#Prep dataframe for Q5 focusing on county level

#Registration Data
county_primary_24 <- read_csv("data/county/county_primary_24.csv")
county_general_24 <- read_csv("data/county/county_general_24.csv")
county_primary_20 <- read_csv("data/county/county_primary_20.csv")
county_primary_20 <- county_primary_20 |>
  rename(County = 1)
county_general_20 <- read_csv("data/county/county_general_20.csv")

#Population Data
county_population <- get_acs(geography="county", 
                       variables = "B01003_001", 
                       state = "MD", year=2022)

#Presidential Election Result (county level)
pres_election_result_county_url <- "https://elections.maryland.gov/elections/2024/general_Results/gen_detail_results_2024_1_1.html"
pres_election_result_county_html <- pres_election_result_county_url |>
  read_html() |>
  html_table()
pres_election_result_county1 <- pres_election_result_county_html[[1]]
pres_election_result_county2 <- pres_election_result_county_html[[2]]
pres_election_result_county2 <- pres_election_result_county2|>
  select(-1)
pres_election_result_county3 <- pres_election_result_county_html[[3]]
pres_election_result_county3 <- pres_election_result_county3|>
  select(-1)
pres_election_result_county4 <- pres_election_result_county_html[[4]]
pres_election_result_county4 <- pres_election_result_county4|>
  select(-1)
pres_election_result_county5 <- pres_election_result_county_html[[5]]
pres_election_result_county5 <- pres_election_result_county5|>
  select(-1)
pres_election_result_county6 <- pres_election_result_county_html[[6]]
pres_election_result_county6 <- pres_election_result_county6|>
  select(-1)
pres_election_result_county <- bind_cols(pres_election_result_county1, pres_election_result_county2, pres_election_result_county3, pres_election_result_county4, pres_election_result_county5, pres_election_result_county6)
```
```{r}
#Primary
#2024
county_primary_24_count <- county_primary_24 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    LIB = sum(LIB),
    OTH = sum(OTH),
    UNA = sum(UNA),
    NLM = sum(NLM)
  ) |>
  mutate(Other = LIB+OTH+NLM) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM24 = 1, REP24 =2, UNA24 = 3, Other24 = 4)
#2020
county_primary_20_count <- county_primary_20 |>
  summarise(
    BAR = sum(BAR),
    DEM = sum(DEM),
    OTH = sum(OTH),
    REP = sum(REP),
    UNA = sum(UNA),
    WCP = sum(WCP)
  ) |>mutate(Other = BAR+OTH+WCP) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM20 = 1, REP20 =2, UNA20 = 3, Other20 = 4)
#Find the change
compare_county_primary_24_20 <- bind_cols(county_primary_24_count, county_primary_20_count) |>
  mutate(DEM_change = ((DEM24-DEM20)/DEM20)*100) |>
  mutate(REP_change = ((REP24-REP20)/REP20)*100) |>
  mutate(UNA_change = ((UNA24-UNA20)/UNA20)*100) |>
  mutate(Other_change = ((Other24-Other20)/Other20)*100)

#General
#2024
county_general_24_count <- county_general_24 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    LIB = sum(LIB),
    OTH = sum(OTH),
    UNA = sum(UNA),
    NLM = sum(NLM),
    GRN = sum(GRN)
  ) |>
  mutate(Other = LIB+OTH+NLM+GRN) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM24 = 1, REP24 =2, UNA24 = 3, Other24 = 4)
#2020
county_general_20_count <- county_general_20 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    BAR = sum(BAR),
    GRN = sum(GRN),
    LIB = sum(LIB),
    WCP = sum(WCP),
    OTH = sum(OTH),
    UNA = sum(UNA)
  ) |>
  mutate(Other = BAR+GRN+LIB+WCP+OTH) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM20 = 1, REP20 =2, UNA20 = 3, Other20 = 4)
#Find the change
compare_county_general_24_20 <- bind_cols(county_general_24_count, county_general_20_count) |>
  mutate(DEM_change = ((DEM24-DEM20)/DEM20)*100) |>
  mutate(REP_change = ((REP24-REP20)/REP20)*100) |>
  mutate(UNA_change = ((UNA24-UNA20)/UNA20)*100) |>
  mutate(Other_change = ((Other24-Other20)/Other20)*100)

#Find Pattern
#Primary24
pattern_county_primary_24 <- county_primary_24 |>
  mutate(Other = LIB+OTH+NLM) |>
  select(County, DEM, REP, UNA, Other, TOTAL) |>
  mutate(DEM_per = (DEM/TOTAL)*100) |>
  mutate(REP_per = (REP/TOTAL)*100) |>
  mutate(UNA_per = (UNA/TOTAL)*100) |>
  mutate(Other_per = (Other/TOTAL)*100) |>
  mutate(DEM_REP = DEM_per-REP_per) |>
  mutate(REP_UNA = REP_per-UNA_per) |>
  mutate(DEM_UNA = DEM_per-UNA_per)
#General24
pattern_county_general_24 <- county_general_24 |>
  mutate(Other = LIB+OTH+NLM+GRN) |>
  select(County, DEM, REP, UNA, Other, TOTAL) |>
  mutate(DEM_per = (DEM/TOTAL)*100) |>
  mutate(REP_per = (REP/TOTAL)*100) |>
  mutate(UNA_per = (UNA/TOTAL)*100) |>
  mutate(Other_per = (Other/TOTAL)*100) |>
  mutate(DEM_REP = DEM_per-REP_per) |>
  mutate(REP_UNA = REP_per-UNA_per) |>
  mutate(DEM_UNA = DEM_per-UNA_per)

#Compare to Population
#Primary24
total_county_primary_24 <- county_primary_24 |>
  select(County, TOTAL) |>
  mutate(County = str_trim(County,side="both"))
cleaned_county_population <- county_population |>
  rename(County = 2, population = 4) |>
  select(2,4) |>
  mutate(County = gsub(" County, Maryland", "", County)) |>
  mutate(County = gsub("Baltimore", "Baltimore County", County)) |>
  mutate(County = gsub("Baltimore County city, Maryland", "Baltimore City", County)) |>
  mutate(County = gsub("St. Mary's", "Saint Mary's", County)) |>
  mutate(County = str_trim(County,side="both"))
regis_population_county_primary_24 <- total_county_primary_24 |>
  left_join(cleaned_county_population, join_by(County)) |>
  mutate(Percent = (TOTAL/population)*100)
#General24
total_county_general_24 <- county_general_24 |>
  select(County, TOTAL) |>
  mutate(County = str_trim(County,side="both"))
regis_population_county_general_24 <- total_county_general_24 |>
  left_join(cleaned_county_population, join_by(County)) |>
  mutate(Percent = (TOTAL/population)*100)

#Compare to the result
#Clean the result
county_result_cleaned <- pres_election_result_county |>
  rename(County = 1, DEM1 = 2, REP1 = 3, LIB = 4, GRN = 5, UNA1 = 6, DEM2 = 7, DEM3 = 8, REP2 = 9, REP3 = 10, OTH1 = 11, OTH2 = 12, OTH3 = 13, OTH4 = 14, OTH5 = 15, OTH6 = 16, UNA2 = 17, UNA3 = 18, UNA4 = 19, UNA5 = 20, UNA6 = 21, UNA7 = 22, UNA8 = 23, UNA9 = 24, UNA10 = 25, UNA11 = 26, UNA12 = 27, UNA13 = 28, UNA14 = 29, UNA15 = 30, UNA16 = 31, UNA17 = 32, UNA18 = 33, UNA19 = 34, OTH7 = 35) |>
  mutate(DEM1 = as.numeric(gsub(",","", DEM1))) |>
  mutate(REP1 = as.numeric(gsub(",","", REP1))) |>
  mutate(LIB = as.numeric(gsub(",","", LIB))) |>
  mutate(GRN = as.numeric(gsub(",","", GRN))) |>
  mutate(UNA1 = as.numeric(gsub(",","", UNA1))) |>
  mutate(DEM2 = as.numeric(gsub(",","", DEM2))) |>
  mutate(DEM3 = as.numeric(gsub(",","", DEM3))) |>
  mutate(REP2 = as.numeric(gsub(",","", REP2))) |>
  mutate(REP3 = as.numeric(gsub(",","", REP3))) |>
  mutate(OTH1 = as.numeric(gsub(",","", OTH1))) |>
  mutate(OTH2 = as.numeric(gsub(",","", OTH2))) |>
  mutate(OTH3 = as.numeric(gsub(",","", OTH3))) |>
  mutate(OTH4 = as.numeric(gsub(",","", OTH4))) |>
  mutate(OTH5 = as.numeric(gsub(",","", OTH5))) |>
  mutate(OTH6 = as.numeric(gsub(",","", OTH6))) |>
  mutate(UNA2 = as.numeric(gsub(",","", UNA2))) |>
  mutate(UNA3 = as.numeric(gsub(",","", UNA3))) |>
  mutate(UNA4 = as.numeric(gsub(",","", UNA4))) |>
  mutate(UNA5 = as.numeric(gsub(",","", UNA5))) |>
  mutate(UNA6 = as.numeric(gsub(",","", UNA6))) |>
  mutate(UNA7 = as.numeric(gsub(",","", UNA7))) |>
  mutate(UNA8 = as.numeric(gsub(",","", UNA8))) |>
  mutate(UNA9 = as.numeric(gsub(",","", UNA9))) |>
  mutate(UNA10 = as.numeric(gsub(",","", UNA10))) |>
  mutate(UNA11 = as.numeric(gsub(",","", UNA11))) |>
  mutate(UNA12 = as.numeric(gsub(",","", UNA12))) |>
  mutate(UNA13 = as.numeric(gsub(",","", UNA13))) |>
  mutate(UNA14 = as.numeric(gsub(",","", UNA14))) |>
  mutate(UNA15 = as.numeric(gsub(",","", UNA15))) |>
  mutate(UNA16 = as.numeric(gsub(",","", UNA16))) |>
  mutate(UNA17 = as.numeric(gsub(",","", UNA17))) |>
  mutate(UNA18 = as.numeric(gsub(",","", UNA18))) |>
  mutate(UNA19 = as.numeric(gsub(",","", UNA19))) |>
  mutate(OTH7 = as.numeric(gsub(",","", OTH7))) |>
  slice(-25) |>
  mutate(DEM = DEM1 + DEM2 + DEM3) |>
  mutate(REP = REP1 + REP2 + REP3) |>
  mutate(UNA = UNA1 + UNA2 + UNA3 + UNA4 + UNA5 + UNA6 + UNA7 + UNA8 + UNA9 + UNA10 + UNA11 + UNA12 + UNA13 + UNA14 + UNA15 + UNA16 + UNA17 + UNA18 + UNA19) |>
  mutate(OTH = OTH1 + OTH2 + OTH3 + OTH4 + OTH5 + OTH6 + OTH7 + LIB + GRN) |>
  select(County, DEM, REP, UNA, OTH) |>
  mutate(Total = DEM+REP+UNA+OTH) |>
  mutate(DEM_per = (DEM/Total)*100) |>
  mutate(REP_per = (REP/Total)*100) |>
  mutate(UNA_per = (UNA/Total)*100) |>
  mutate(OTH_per = (OTH/Total)*100) |>
  mutate(DEM_REP = DEM_per-REP_per) |>
  mutate(REP_UNA = REP_per-UNA_per) |>
  mutate(DEM_UNA = DEM_per-UNA_per)
```
```{r}
#Compare Presidential result between each candidate.
CANDIDATES_county_result_cleaned <- pres_election_result_county |>
  rename(County = 1, DEM1 = 2, REP1 = 3, LIB = 4, GRN = 5, UNA1 = 6, DEM2 = 7, DEM3 = 8, REP2 = 9, REP3 = 10, OTH1 = 11, OTH2 = 12, OTH3 = 13, OTH4 = 14, OTH5 = 15, OTH6 = 16, UNA2 = 17, UNA3 = 18, UNA4 = 19, UNA5 = 20, UNA6 = 21, UNA7 = 22, UNA8 = 23, UNA9 = 24, UNA10 = 25, UNA11 = 26, UNA12 = 27, UNA13 = 28, UNA14 = 29, UNA15 = 30, UNA16 = 31, UNA17 = 32, UNA18 = 33, UNA19 = 34, OTH7 = 35) |>
  mutate(DEM1 = as.numeric(gsub(",","", DEM1))) |>
  mutate(REP1 = as.numeric(gsub(",","", REP1))) |>
  mutate(LIB = as.numeric(gsub(",","", LIB))) |>
  mutate(GRN = as.numeric(gsub(",","", GRN))) |>
  mutate(UNA1 = as.numeric(gsub(",","", UNA1))) |>
  mutate(DEM2 = as.numeric(gsub(",","", DEM2))) |>
  mutate(DEM3 = as.numeric(gsub(",","", DEM3))) |>
  mutate(REP2 = as.numeric(gsub(",","", REP2))) |>
  mutate(REP3 = as.numeric(gsub(",","", REP3))) |>
  mutate(OTH1 = as.numeric(gsub(",","", OTH1))) |>
  mutate(OTH2 = as.numeric(gsub(",","", OTH2))) |>
  mutate(OTH3 = as.numeric(gsub(",","", OTH3))) |>
  mutate(OTH4 = as.numeric(gsub(",","", OTH4))) |>
  mutate(OTH5 = as.numeric(gsub(",","", OTH5))) |>
  mutate(OTH6 = as.numeric(gsub(",","", OTH6))) |>
  mutate(UNA2 = as.numeric(gsub(",","", UNA2))) |>
  mutate(UNA3 = as.numeric(gsub(",","", UNA3))) |>
  mutate(UNA4 = as.numeric(gsub(",","", UNA4))) |>
  mutate(UNA5 = as.numeric(gsub(",","", UNA5))) |>
  mutate(UNA6 = as.numeric(gsub(",","", UNA6))) |>
  mutate(UNA7 = as.numeric(gsub(",","", UNA7))) |>
  mutate(UNA8 = as.numeric(gsub(",","", UNA8))) |>
  mutate(UNA9 = as.numeric(gsub(",","", UNA9))) |>
  mutate(UNA10 = as.numeric(gsub(",","", UNA10))) |>
  mutate(UNA11 = as.numeric(gsub(",","", UNA11))) |>
  mutate(UNA12 = as.numeric(gsub(",","", UNA12))) |>
  mutate(UNA13 = as.numeric(gsub(",","", UNA13))) |>
  mutate(UNA14 = as.numeric(gsub(",","", UNA14))) |>
  mutate(UNA15 = as.numeric(gsub(",","", UNA15))) |>
  mutate(UNA16 = as.numeric(gsub(",","", UNA16))) |>
  mutate(UNA17 = as.numeric(gsub(",","", UNA17))) |>
  mutate(UNA18 = as.numeric(gsub(",","", UNA18))) |>
  mutate(UNA19 = as.numeric(gsub(",","", UNA19))) |>
  mutate(OTH7 = as.numeric(gsub(",","", OTH7))) |>
  slice(-25) |>
  mutate(Total = DEM1+REP1+LIB+GRN+UNA1+DEM2+DEM3+REP2+REP3+OTH1+OTH2+OTH3+OTH4+OTH5+OTH6+UNA2+UNA3+UNA4+UNA5+UNA6+UNA7+UNA8+UNA9+UNA10+UNA11+UNA12+UNA13+UNA14+UNA15+UNA16+UNA17+UNA18+UNA19+OTH7) |>
  mutate(DEM1per = (DEM1/Total)*100) |>
  mutate(REP1per = (REP1/Total)*100) |>
  mutate(LIBper = (LIB/Total)*100) |>
  mutate(GRNper = (GRN/Total)*100) |>
  mutate(UNA1per = (UNA1/Total)*100) |>
  mutate(DEM2per = (DEM2/Total)*100) |>
  mutate(DEM3per = (DEM3/Total)*100) |>
  mutate(REP2per = (REP2/Total)*100) |>
  mutate(REP3per = (REP3/Total)*100) |>
  mutate(OTH1per = (OTH1/Total)*100) |>
  mutate(OTH2per = (OTH2/Total)*100) |>
  mutate(OTH3per = (OTH3/Total)*100) |>
  mutate(OTH4per = (OTH4/Total)*100) |>
  mutate(OTH5per = (OTH5/Total)*100) |>
  mutate(OTH6per = (OTH6/Total)*100) |>
  mutate(UNA2per = (UNA2/Total)*100) |>
  mutate(UNA3per = (UNA3/Total)*100) |>
  mutate(UNA4per = (UNA4/Total)*100) |>
  mutate(UNA5per = (UNA5/Total)*100) |>
  mutate(UNA6per = (UNA6/Total)*100) |>
  mutate(UNA7per = (UNA7/Total)*100) |>
  mutate(UNA8per = (UNA8/Total)*100) |>
  mutate(UNA9per = (UNA9/Total)*100) |>
  mutate(UNA10per = (UNA10/Total)*100) |>
  mutate(UNA11per = (UNA11/Total)*100) |>
  mutate(UNA12per = (UNA12/Total)*100) |>
  mutate(UNA13per = (UNA13/Total)*100) |>
  mutate(UNA14per = (UNA14/Total)*100) |>
  mutate(UNA15per = (UNA15/Total)*100) |>
  mutate(UNA16per = (UNA16/Total)*100) |>
  mutate(UNA17per = (UNA17/Total)*100) |>
  mutate(UNA18per = (UNA18/Total)*100) |>
  mutate(UNA19per = (UNA19/Total)*100) |>
  mutate(OTH7per = (OTH7/Total)*100) |>
  select(-2,-3,-4,-5,-6,-7,-8,-9,-10,-11,-12,-13,-14,-15,-16,-17,-18,-19,-20,-21,-22,-23,-24,-25,-26,-27,-28,-29,-30,-31,-32,-33,-34,-35)
```
```{r}
#Compare Presidential race results (only main candidates) from 2024, 2020 and 2016
#2024
MAIN_CANDIDATES_2024 <- pres_election_result_county |>
  select(1,2,3) |>
  rename(Harris = 2, Trump = 3) |>
  mutate(Harris = as.numeric(gsub(",","", Harris))) |>
  mutate(Trump = as.numeric(gsub(",","", Trump))) |>
  mutate(Total = Harris + Trump) |>
  mutate(Harris2024 = (Harris/Total)*100) |>
  mutate(Trump2024 = (Trump/Total)*100) |>
  select(1,5,6)
#2020
pres_election_result_county_2020_url <- "https://elections.maryland.gov/elections/2020/results/general/gen_detail_results_2020_4_BOT001-.html"
pres_election_result_county_2020_html <- pres_election_result_county_2020_url |>
  read_html() |>
  html_table()
pres_election_result_2020_county <- pres_election_result_county_2020_html[[2]]
MAIN_CANDIDATES_2020 <- pres_election_result_2020_county |>
  select(1,2,3) |>
  rename(Trump = 2, Biden = 3) |>
  mutate(Trump = as.numeric(gsub(",","", Trump))) |>
  mutate(Biden = as.numeric(gsub(",","", Biden))) |>
  mutate(Total = Trump + Biden) |>
  mutate(Trump2020 = (Trump/Total)*100) |>
  mutate(Biden2020 = (Biden/Total)*100) |>
  select(6,5)
#2016
pres_election_result_county_2016_url <- "https://elections.maryland.gov/elections/2016/results/general/gen_detail_results_2016_4_BOT001-.html"
pres_election_result_county_2016_html <- pres_election_result_county_2016_url |>
  read_html() |>
  html_table()
pres_election_result_2016_county <- pres_election_result_county_2016_html[[1]]
MAIN_CANDIDATES_2016 <- pres_election_result_2016_county |>
  select(1,2,3) |>
  rename(Trump = 2, Clinton = 3) |>
  mutate(Trump = as.numeric(gsub(",","", Trump))) |>
  mutate(Clinton = as.numeric(gsub(",","", Clinton))) |>
  mutate(Total = Trump + Clinton) |>
  mutate(Trump2016 = (Trump/Total)*100) |>
  mutate(Clinton2016 = (Clinton/Total)*100) |>
  select(6,5)
MAIN_CANDIDATES_COMPARE <- bind_cols(MAIN_CANDIDATES_2024,MAIN_CANDIDATES_2020,MAIN_CANDIDATES_2016)
MAIN_CANDIDATES_COMPARE <- MAIN_CANDIDATES_COMPARE |>
  select(1,3,5,7)
```

Thank you Mr. Willis, have a great winter break.❄️😊