---
title: "lab_06"
author: "Molecule Jongwilai"
date: "2024-10-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

-   Tabula

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse and any other libraries you'll need
library(tidyverse)

```

## Get Our PDF

We'll be working with the [Maryland monthly voter registration update](https://elections.maryland.gov/pdf/vrar/2024_09.pdf) from the State Board of Elections. You'll want to download it to a place you'll remember (like your Downloads folder, or the labs folder in your repository). The goal is to write a couple of paragraphs that summarize the changes in voter registration in Maryland during the month of September.

## Setup Tabula

Start Tabula, then go to <http://127.0.0.1:8080/> in your browser. Click the "Browse" button and find the PDF file and click "open", and then click the "Import button" in Tabula. This will take a few seconds.

This PDF has multiple possible tables to extract. We're going to make four dataframes: 

* new registrations by party
* removals by party
* changes in registration
* current registrations by county.

You will need to draw boxes around four sections and export four CSV files that you will load into R. In Tabula, draw a box around the borders of those tables and click the "Preview & Export Extracted Data" button for each one. The export should look pretty clean, but might be missing some column headers. You'll need to include total columns.

Export each CSV (each one should be called `tabula-2024-09.csv` by default, so you should rename them for clarity) to your lab_06/data folder.

From there, you will need to read in and clean up the data, and in particular the headers. You can choose to include the headers from the PDF in your exported CSV files OR to exclude them and add them when importing. The pre_lab has an example of how to do this!

## Load and clean up the data in R

You will need to read in and clean up the data so that it can be used for analysis. By "clean" I mean the column headers must not contain spaces and they should have meaningful names, not "x1" or something similar. The current registrations dataframe *MUST* include the county name, but should not include the columns under `Changes` and `Party Affiliation From`. How you do that is up to you (and could involve something outside R), but you can use `select()` with or without the minus sign to include or exclude certain columns. You also can use the `rename` function to, well, rename columns.

```{r}
# Party registrations
new_reg_2024 <- read_csv("data/2024/new_reg_by_party_2024_09.csv") |>
  rename(METHOD_OF_REGISTRATION = ...1) 
view(new_reg_2024)
```

```{r}
# Removals
removals_2024 <- read_csv("data/2024/removals_by_party_2024_09.csv") |>
  rename(MEAN_OR_REASON_OF_REMOVAL = ...1) |>
  filter(!is.na(MEAN_OR_REASON_OF_REMOVAL))
view(removals_2024)
```

```{r}
# Changes
changes_in_reg_2024 <- read_csv("data/2024/changes_in_reg_2024_09.csv") |>
  rename(COUNTY = ...1)
view(changes_in_reg_2024)
```

```{r}
# Current registrations
current_reg_2024 <- read_csv("data/2024/current_reg_by_county_2024_09.csv", col_name = FALSE) |>
  select(X1, X12, X13, X14, X15, X16, X17, X18, X19) |>
  rename(COUNTY = X1, DEM = X12, REP = X13, NLM = X14, LIB = X15, GRN = X16, UNAF = X17, OTH = X18, TOTAL = X19)
view(current_reg_2024)
```

## Rinse, Repeat: September 2022 data

Do all of that again, for the [September 2022 voter registration update](https://elections.maryland.gov/pdf/vrar/2022_09.pdf), so we can compare the two. If you've done well, this is easy: just do the same steps in Tabula, export the files and use the same code to read in and clean up the data, *just remember to give these dataframes different names* so that you can compare September 2024 to September 2022.

To answer the following questions, you'll need to calculate percentages and do other calculations. Whether you do that by combining or joining dataframes or you write code to run separately on both is up to you.

```{r}
# Party registrations
new_reg_2022 <- read_csv("data/2022/new_reg_by_party_2022_09.csv") |>
  rename(METHOD_OF_REGISTRATION = ...1) 
view(new_reg_2022)
```

```{r}
# Removals
removals_2022 <- read_csv("data/2022/removals_by_party_2022_09.csv") |>
  rename(MEAN_OR_REASON_OF_REMOVAL = ...1) |>
  filter(!is.na(MEAN_OR_REASON_OF_REMOVAL))
view(removals_2022)
```

```{r}
# Changes
changes_in_reg_2022 <- read_csv("data/2022/changes_in_reg_2022_09.csv") |>
  rename(COUNTY = ...1)
view(changes_in_reg_2022)
```

```{r}
# Current registrations
current_reg_2022 <- read_csv("data/2022/current_reg_by_county_2022_09.csv", col_name = FALSE) |>
  select(X1, X12, X13, X14, X15, X16, X17, X18, X19) |>
  rename(COUNTY = X1, DEM = X12, REP = X13, GRN = X14, LIB = X15, WCP = X16, UNAF = X17, OTH = X18, TOTAL = X19)
view(current_reg_2022)
```

## Answer questions

Q1. Calculate the percentage of registered Democrats, Republicans and unaffiliated voters for the entire state in September 2024 and September 2022, and then calculate the percentage change for those three party categories (the formula for percentage change is easy to remember: (New-Old)/Old). 

How would you describe the changes in these partisan categories in a story? 

What's the most interesting thing about these changes that would be the lede of that story?

A1.
- I would describe that comparing Maryland voters in Sept. 2024 to those in Sept. 2022, Democratic voters slightly decreased by about 1.40%; on the other hand, Republican voters slightly increased by about 1.04%. However, most notably, voters in Maryland who identify as Unaffiliated increased by almost 10%.

- I think the most interesting of these findings is the increasing number of people who declare to be Unaffiliated. We could find out more about why many people in Maryland decided to register as unaffiliated voters this year.    

```{r}
current_reg_2024 |>
  filter(COUNTY == "TOTAL")
#DEM
(2219417/4231403)*100
#REP
(1009323/4231403)*100
#UNAF
(923927/4231403)*100
```
```{r}
current_reg_2022 |>
  filter(COUNTY == "TOTAL")
#DEM
(2250961/4161637)*100
#REP
(998905/4161637)*100
#UNAF
(841573/4161637)*100
```
```{r}
#Percent Change
#DEM
((2219417 - 2250961)/2250961)*100
#REP
((1009323 - 998905)/998905)*100
#UNAF
((923927 - 841573)/841573)*100
```
Q2. Among the Maryland jurisdictions with the most registered voters, which one had the biggest percentage increase in unaffiliated voters between 2022 and 2024? 

How did that county vote for governor in 2022 and for president in 2020? 

Contrast the partisan percentages of that county with the county that had the eighth-largest percentage increase in unaffiliated voters. 

You should display information from the data that shows the counties with the largest percentage increase first.

A2.
- Prince George's County had the biggest percentage increase of unaffiliated voters with about 17.20% from 2022 to 2024.

- According to the Maryland State Board of Election's website (https://elections.maryland.gov/elections/2022/general_results/gen_detail_results_2022_1_1.html, https://elections.maryland.gov/elections/2020/results/general/gen_detail_results_2020_4_BOT001-.html), in 2022, most voters in Prince George's County voted Wes Moore from Democrat for the governor, and in 2020, most voters voted Joe Biden from Democrat for the president. 

- The county that had the eighth-largest percentage increase for unaffiliated voters is Frederick County. In 2024, Prince George's County had the third-smallest percentage of unaffiliated voters with 16.48% compared to Frederick County which had the largest percentage with 26.44%. In 2022, Prince George's County had the second-smallest percentage of unaffiliated voters with 14.30% compared to Frederick County which also had the largest percentage with 24.80%.

```{r}
#Find County that has the biggest percentage increase in unaffiliated voters between 2022 and 2024.
reg_unaf_2024 <- current_reg_2024 |>
  select(COUNTY, UNAF) |>
  rename(UNAF_24 = UNAF)
reg_unaf_2022 <- current_reg_2022 |>
  select(COUNTY, UNAF) |>
  rename(UNAF_22 = UNAF)
reg_unaf_changes <- reg_unaf_2024 |>
  full_join(reg_unaf_2022, join_by(COUNTY))
reg_unaf_changes |>
  mutate(PERCENT_CHANGE = ((UNAF_24-UNAF_22)/UNAF_22)*100) |>
  arrange(desc(PERCENT_CHANGE))

#Find the percentage of unaffiliated voters in Prince George's and Frederick from 2022 and 2024.
current_reg_2024 |>
  select(COUNTY, UNAF, TOTAL) |>
  mutate(PERCENTAGE = (UNAF/TOTAL)*100) |>
  arrange(desc(PERCENTAGE))
current_reg_2022 |>
  select(COUNTY, UNAF, TOTAL) |>
  mutate(PERCENTAGE = (UNAF/TOTAL)*100) |>
  arrange(desc(PERCENTAGE))
```

Q3. Do the party percentages of new registrants (% Dem, % Rep, % Unaffiliated) from September 2024 and September 2022 match the overall percentage of partisan registration in the state at those times?

In other words, did new registrants mirror the overall partisan composition of the electorate, or were they different?

Do any of the methods of registration vary significantly from the overall partisan pattern? Which method is worthy of further exploration?

A3.
- The party percentages of new registrants are not absolutely in line with the overall percentage of partisan registration in the state both in 2024 and 2022. In 2024, the percentage of Republican registration (23.85%) is larger than that of Unaffiliated registration (21.48%); on the contrary, the party percentage of new registrants for Republican (18.23%) is smaller than that for Unaffiliated (28.68%). Similarly, In 2022, the percentage of Republican registration (24.00) is larger than that of Unaffiliated registration (20.22%); on the contrary, the party percentage of new registrants for Republican (15.05%) is smaller than that for Unaffiliated (30.60%).   

- I think the method that is interesting and worthy of more investigation is in-person registration, especially in 2024, the percentage of new voters using this method for Republican registration (26.87%) is larger than that for Unaffiliated registration (21.45%) while overall, the percentage of new voters from every method for Republican (18.23%) is smaller than those for Unaffiliated (28.68%).   

```{r}
#2024
#Percentage of Partisan Reg is DEM 52.45%, REP 23.85%, UNAF 21.84%
new_reg_2024 |>
  select(METHOD_OF_REGISTRATION, DEM, REP, UNAF, TOTAL) |>
  mutate(DEM_PERCENT = ((DEM/TOTAL)*100), REP_PERCENT = ((REP/TOTAL)*100), UNAF_PERCENT = ((UNAF/TOTAL)*100))
#Percentage of New Reg is DEM 49.85%, REP 18.23%, UNAF 28.68%
#Percentage of In Person Method is DEM 48.58%, REP 26.87%, UNAF 21.45%
```
```{r}
#2022
#Percentage of Partisan Reg is DEM 54.09%, REP 24.00%, UNAF 20.22%
new_reg_2022 |>
  select(METHOD_OF_REGISTRATION, DEM, REP, UNAF, TOTAL) |>
  mutate(DEM_PERCENT = ((DEM/TOTAL)*100), REP_PERCENT = ((REP/TOTAL)*100), UNAF_PERCENT = ((UNAF/TOTAL)*100))
#Percentage of New Reg is DEM 51.64%, REP 15.05%, UNAF 30.60%
#Percentage of In Person Method is DEM 52.22%, REP 21.11%, UNAF 23.33%
```
Q4. Choose your own adventure! Using both the 2024 and 2022 data for one of the dataframes (changes, removals or new registrations), ask and answer a question that makes at least one comparison and requires code to answer. That could involve calculating a percentage, percentage change or the raw difference between two columns.

A4.
- In 2024 and 2022, what are the means or reasons that voters use or cause voters to remove registration from all parties the most? 
# Transfering out of County is the reason that caused the largest number of voters to remove registration from all parties in both 2024 and 2022. 

- In 2024 and 2022, what is the percentage of voters who remove registration from Democrat, Republican, and Unaffiliated compared to all of the voters in the removal record?
# In 2024 the percentage of voters who remove the registration from Democrat, Republican, and Unaffiliated is 57.99%, 19.18%, and 20.83% respectively. In 2022 the percentage of voters who remove registration from Democrat, Republican, and Unaffiliated is 57.78%, 18.88%, and 21.35% respectively.  

- What is the percentage change from 2022 to 2024 of voters who remove registration from Democrat, Republican, and Unaffiliated? 
# The percentage change from 2022 to 2024 of voters who remove the registration from Democrat, Republican, and Unaffiliated decreased by 16.02%, 15.08%, and 18.35% respectively.  
```{r}
#Find the the mean or reason that voters use or cause voter to remove registration the most.
#2024
removals_2024 |>
  arrange(desc(TOTAL))
#2022
removals_2022 |>
  arrange(desc(TOTAL))

#Find the percentage of voters who remove registration from Democrat, Republican and Unaffiliated.
#2024
removals_2024 |>
  select(MEAN_OR_REASON_OF_REMOVAL, DEM, REP, UNAF, TOTAL) |>
  filter(MEAN_OR_REASON_OF_REMOVAL == "TOTAL") |>
  mutate(DEM_PERCENT = ((DEM/TOTAL)*100), REP_PERCENT = ((REP/TOTAL)*100), UNAF_PERCENT = ((UNAF/TOTAL)*100))
#2022
removals_2022 |>
  select(MEAN_OR_REASON_OF_REMOVAL, DEM, REP, UNAF, TOTAL) |>
  filter(MEAN_OR_REASON_OF_REMOVAL == "TOTAL") |>
  mutate(DEM_PERCENT = ((DEM/TOTAL)*100), REP_PERCENT = ((REP/TOTAL)*100), UNAF_PERCENT = ((UNAF/TOTAL)*100))

#Find the percentage change from 2022 to 2024 of voter who remove registration from Democrat, Republican and Unaffiliated.
removals_2024_for_join <- removals_2024 |>
  select(MEAN_OR_REASON_OF_REMOVAL, DEM, REP, UNAF) |>
  filter(MEAN_OR_REASON_OF_REMOVAL == "TOTAL") |>
  rename(DEM_24 = DEM, REP_24 = REP, UNAF_24 = UNAF)
removals_2022_for_join <- removals_2022 |>
  select(MEAN_OR_REASON_OF_REMOVAL, DEM, REP, UNAF) |>
  filter(MEAN_OR_REASON_OF_REMOVAL == "TOTAL") |>
  rename(DEM_22 = DEM, REP_22 = REP, UNAF_22 = UNAF)
removal_change <- removals_2024_for_join |>
  full_join(removals_2022_for_join, join_by(MEAN_OR_REASON_OF_REMOVAL))
removal_change |>
  mutate(DEM_CHANGE = ((DEM_24-DEM_22)/DEM_22)*100, REP_CHANGE = ((REP_24-REP_22)/REP_22)*100, UNAF_CHANGE = ((UNAF_24-UNAF_22)/UNAF_22)*100) 
```

Q5. What's the best story idea or question you've seen as a result of the work you've done in this lab?

A5. I think people who register as unaffiliated voters in Maryland could be an interesting idea for the story;
- We could explore more on why the number of people who declare as unaffiliated voters increased by almost 10% from September 2022 to September 2024. 
- We could investigate more on why Prince George's County had the most percentage increase of unaffiliated voters, and why Frederick County had the most unaffiliated voters in both 2022 and 2024.
- We could dive into new registration data to see why the percentage of unaffiliated voters is larger than those of Republican voters both in 2022 and 2024.   
