---
title: "lab_01"
author: "Molecule Jongwilai"
date: "2024-09-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**). When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the slice of WinRed contribution data from West Virginia, which is in a file called `lab_01.rds`

**Task** Using the example from the pre_lab, write code to load the RDS file and assign it to a good variable name.

```{r}
west_verginia_winred_contributions <- read_rds("lab_01.rds")
```

## Examine Data

**Task** Run the codeblock below to use glimpse to get a sense of the column names, types and sample data.

**Q1** How many rows and columns does this dataframe have?

**A1**
Rows = 40,441
Columns = 24

```{r}
glimpse(west_verginia_winred_contributions)
```

## Answer questions

**Q2** Which West Virginia city provided the largest number of contributions? How many?

**A2** Charleston is a city that has the largest number of contributions which is 3226. 

```{r}
west_verginia_winred_contributions |>
  group_by(city) |>
  summarise(
    count_contributions = n()
  ) |>
  arrange(desc(count_contributions))
```

**Q3** Which West Virginia cities accounted for the fewest contributions? How many of them, roughly speaking?

**A3** There are 63 cities, Allen Junction, Alum Bridge, BRIDGEPORT, Bergoo, Berkeley Spgs, Bloomingrose, Breeden, Brenton, Bristol, Buckeye, Canvas, Ccbxxthvc, Clothier, Coal city, Costa, DANIELS, Dawes, Dingess, Dunmore, Durbin, Exchange, Falling Wtrs, Flat Top, Folsom, Gallagher, Griffithsville, Harman, Harper, Hines, Hometown, Hundred, Ireland, Lochgelly, Lorado, Marmet, Meador, Moatsville, Nebo, Newburg, North Matewan, Omar WV, Orgas, Palestine, Peytona, Ragland, Ranger, Rippon, S Charleston, Secondcreek, Shock, Slab Fork, Squire, St Albans, Tallmansville, Valley Fork, Walkersville, Wana, Wayside, West Liberty, West Ville, Winifrede, Wolfcreek, milton, accounted for the fewest contribution which is one contribution from each city. 

```{r}
west_verginia_winred_contributions |>
  group_by(city) |>
  summarise(
    count_contributions = n()
  ) |>
  arrange(count_contributions)
```

**Q4** What's your best educated guess to explain why the city identified in question 2 had more contributions compared to the cities identified in question 3? Hint: you won't find the answer in the data, you'll need to do some quick web research. You do not need to write any code for this question, so no codeblock is provided.

**A4** I surmise that this is mainly because Charleston is the capital of West Virginia, and according to census.gov, its population is 46,838, compared to cities in question number 3 like Bridgeport or St. Albans, which have fewer populations of 9,377 and 10,387, respectively. Therefore, there are more contributions in Charleston than in Bridgeport or St. Albans.   

**Q5** What recipient got the largest total amount of money? How many contributions did they receive and what was the average donation?

**A5** TRUMP NATIONAL COMMITTEE JFC, INC. received the largest total amount, with 14,063 contributions and an average donation of $32.9904288. 

```{r}
west_verginia_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    total_amount = sum(amount),
    memo_text_count_contributions = n(),
    average_amount = mean(amount)
  ) |>
  arrange(desc(total_amount)) 
```

**Q6** What is that top recipient and what explains its popularity? Hint: this will require web research, you won't be able to answer this purely from the data set. You do not need to write any code for this question, so no codeblock is provided.

**A6** The top recipient is TRUMP NATIONAL COMMITTEE JFC, INC., and according to the Federal Election Commission website (https://docquery.fec.gov/cgi-bin/forms/C00828541/1777183/), it is a part of the Republican Party's principal campaign, DONALD J. TRUMP FOR PRESIDENT 2024, INC. Hence, I think this can explain its popularity. 

In addition, if you search keyword 'Trump Donation' on Google, this website, https://win.donaldjtrump.com/trump-national-committee-jfc/vp-convention-gs/?utm_medium=ads&utm_source=google_lp_kw&utm_campaign=20240718_tnc_search-trial_google_search_lp&utm_content=donate&utm_term=donate&gad_source=1&gclid=Cj0KCQjw8--2BhCHARIsAF_w1gzRdkqXRsaV5Tby75AVIMlwHepZVRWWFp1Cj3dbEpTmnLpP8j2Suc8aAo2dEALw_wcB, will be the first link that comes up which if you donate, it will be wired to Trump National Committee JFC. Therefore, I think it's the top recipient because it's the easiest destination to find when Republican supporters want to donate.     

**Q7**. Using the answer from A5 as a starting point, produce a summary table that has the following information for each recipient: \* count of contributions \* total amount of all contributions \* the average contribution amount \* the median contribution amount \* the smallest contribution amount \* the largest contribution amount

Store it as a variable object called `wv_summary` so that you can examine it with the data viewer. What newsworthy nuggets jump out at you that you could imagine dropping into a news story? Write a few summary sentences with overall findings.

**A7**. The summary shows that TRUMP NATIONAL COMMITTEE is the top recipient with the largest number of contributions, 14,063, and the highest amount of money, $463,944.40, in the West Virginia area. Furthermore, data also shows that the average amount of donated money, $32.9904288, is higher than the median amount, $20.240, suggesting that most donors gave a smaller amount of money for this campaign. 

```{r}
wv_summary <- west_verginia_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    memo_text_count_contributions = n(),
    total_amount = sum(amount),
    average_amount = mean(amount),
    median_amount = median(amount),
    min_amount = min(amount),
    max_amount = max(amount)
  ) |>
  arrange(desc(total_amount)) 
```