---
title: "lab_07"
author: "Molecule Jongwilai"
date: "2024-10-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

-   A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
library(tidycensus)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key *only* if you haven't already

```{r echo=FALSE}
# census_api_key("YOUR API KEY", overwrite=TRUE)
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2022 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2022 5-year ACS and save them to a dataframe

```{r}
acs_2022_5y <- load_variables(
  2022, dataset = "acs5", cache = FALSE)
```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2022 5-year estimates? You will need to find the variable for median age first.

**A1** Baltimore City is the jurisdiction in Maryland that have the lowest median age at 36.0.  

```{r}
maryland_med_age <- get_acs(geography = "county",
              variables = c(med_age = "B01002_001"),
              state = "MD",
              year = 2022)

maryland_med_age |>
  arrange(estimate)
```

**Q2** Which Maryland counties have a median age of at least 50? You can use the dataframe you just made to answer this, but you must write code to display only those counties with a median age of at least 50.

**A2** Talbot and Worcester are the two counties that have media age of at least 50. Talbot's median age is 54.4, and Worcester's median age is 50.6.   

```{r}
maryland_med_age |>
  filter(estimate > "50.00")
```

## Ticket-Splitting

Ticket-splitting (voting for candidates from multiple parties) in today's political environment seems like a rarity, but it does happen, even in states like Maryland. For the remaining questions you'll be looking at cast vote records from Frederick County - Maryland's most competitive county - for the 2022 general election. You'll be trying to identify the zip codes where ticket-splitting is happening the most and to learn more about those places. To do that, you'll need to read in two CSV files, join them and then also bring in data from the Census API.

### Setup

Load the two CSV files in the `data` folder: one containing the ballot choices of Frederick voters in five races (governor, comptroller, attorney general and U.S. House) and the other containing information about Frederick's voting locations. You'll need to join them together.

```{r}
frederick_ballot_choices <- read_csv("data/frederick_cvr.csv")
frederick_voting_locations <- read_csv("data/frederick_precincts.csv")
frederick_ballot_and_locations <- frederick_ballot_choices |> 
  left_join(frederick_voting_locations, join_by(precinct))
```

**Q3** Let's answer some basic questions about ticket-splitting: 
- how many voters voted for different parties in the governor and comptroller's race? 
- How many voted for the same party in both races? 
- What percentage of all voters split their tickets in those races? 
- In how many cases did voters choose the same party in all five contests? 
- What percentage of all voters does that represent?

**A3**
- There are 11,871 voters that voted for different parties in the governor and comptroller's race.
- There are 95,221 voters that voted for the same party in the governor and comptroller's race.
- The percentage of voters who split their tickets in the governor and comptroller's race is 11.06%. 
- There are 89,916 voters that voted for the same party in in all five contests.
- The percentage of voters who voted for the same party in in all five contests is 83.80%. 

```{r}
frederick_ballot_and_locations |>
  filter(governor_lt_governor != comptroller) |>
  summarise(
    count = n()
  )
frederick_ballot_and_locations |>
  filter(governor_lt_governor == comptroller) |>
  summarise(
    count = n()
) 
#Percentage of all ticket splitter
(11871/107302)*100
frederick_ballot_and_locations |>
  filter(governor_lt_governor == comptroller & comptroller == attorney_general & attorney_general == u_s_senator & u_s_senator == representative_in_congress_6) |>
  summarise(
    count = n()
) 
#Percentage of voters who choose the same party in all five races
(89916/107302)*100
```

**Q4** Now let's look for the zip codes where voters preferred Republicans in every race *except* the governor's race, where Dan Cox lost to Wes Moore by a wide margin. 

Write code to generate a dataframe of zip codes with the number of those voters for each of those zip codes, ordering the dataframe so that the zip code with the most ticket-splitters appears first.

**A4** 21702, 21701, 21793, 21703, and 21771 are the zip codes of areas that have the top-five number of voters who preferred Republicans in every race except the governor's race as most voters in these areas voted for Wes Moore instead of Dan Cox.

```{r}
frederick_ballot_and_zipcodes <- frederick_ballot_and_locations |>
  filter(governor_lt_governor == "DEM", comptroller == "REP",  attorney_general == "REP", u_s_senator == "REP", representative_in_congress_6 == "REP") |>
  group_by(zipcode) |> 
  summarise(
    count_zip = n()
  ) |>
  arrange(desc(count_zip))
```

**Q5** Let's find out more about those zip codes from A4. Using tidycensus, get a dataframe of Maryland zip codes and their median ages, then join it to the dataframe you produced in A4. 

NOTE: you'll need to ensure that the two columns you use for the join are the same datatype. 

Summarize your findings; how would you describe the zip codes with more non-Cox Republicans compared to those with fewer? Where are those leading zip codes located?

**A5** 
The zip codes with more non-Cox Republicans voters e.g. 21702, 21701, 21793, 21703, 21771, etc.: 
- tend to have a slightly younger population with a median age spanning from 35.1 to 42.1 compared to areas that have fewer non-Cox Republican voters.
- and according to censusreporter.org, these areas tend to have a huge population spanning from 10,882 to 49,596 compared to areas that have fewer non-Cox Republican voters.

The zip codes with fewer non-Cox Republicans voters e.g. 21780, 21718, 21777, 21791, 21714, etc.: 
- tend to have a slightly older population with a median age spanning from 36.9 to 46.6 compared to areas that have more non-Cox Republican voters.
- and according to censusreporter.org, these areas tend to have a small population spanning from 143 to 4,855 compared to areas that have more non-Cox Republican voters.

```{r}
md_zipcodes_med_age <- get_acs(geography="zcta", 
                       variables = "B01002_001", 
                       state='MD', year=2019)

glimpse(frederick_ballot_and_zipcodes)
glimpse(md_zipcodes_med_age)

md_zipcodes_med_age <- md_zipcodes_med_age |>
  mutate(GEOID = as.numeric(GEOID))
glimpse(md_zipcodes_med_age)

frederick_zipcodes_med_age <- frederick_ballot_and_zipcodes |>
  left_join(md_zipcodes_med_age, join_by(zipcode==GEOID))

frederick_zipcodes_med_age
```