---
title: "lab_10"
author: "Molecule Jongwilai"
date: "2024-11-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

-   the tidyverse, janitor and tigris libraries
-   you might choose to use the tidycensus library for Q3

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
install.packages('ggalt')
library(ggalt)
```
```{r}
# Turn off scientific notation
options(scipen=999)

# Load libraries
library(tidyverse)
library(sf)
library(janitor)
library(tidycensus)
library(ggalt)
```

Let's explore the election results from last night and make some maps!

## Questions

**Q1.** Make a county-level map of Maryland's results for U.S. Senate using `md_senate_county_24.csv` in the data folder, calculating the difference between Angela Alsobrooks's percentage of the total votes and Larry Hogan's percentage of the total votes and displaying that percentage difference on the map. What is the story here, and what is the county with the most interesting results?

**A1.** I think the story idea here is the data shows that Alsobrooks obtained more votes than Hogan, especially in counties in the central part of Maryland. We might investigate the demographics of voters in these areas more and how they differ from other Maryland areas where Hogan gained more votes. The most interesting counties to start might be Prince George's and Baltimore City where Alsobrooks won with more than a 50% percentage difference compared to Hogan's votes.   

```{r}
#Download Election Result 
md_senate_county_24 <- read_csv("data/md_senate_county_24.csv")

#Download County Map 
md_county_map <- get_acs(geography = "county",
              variables = c(population = "B01001_001"),
              state = "MD",
              year = 2022,
              geometry = TRUE)

#Clean Map data 
md_county_map <- md_county_map |>
  rename(County = NAME) |>
  mutate(County = str_to_title(str_remove_all(County,", Maryland|County"))) |>
  mutate(County = str_trim(County,side="both")) |>
  select(County, geometry) |>
  mutate(County = if_else(County == "St. Mary's", "Saint Mary's", County)) |> 
  mutate(County = if_else(County == "Baltimore", "Baltimore County", County))

#Join Senate Data with Map
md_senate_county_24_with_map <- md_county_map |>
  #make sure the dataframe with geometry come first otherwise you will get error when you make map 
  inner_join(md_senate_county_24, join_by(County))

#Find the Difference Percentage
md_senate_county_24_with_map <- md_senate_county_24_with_map |>
  mutate(total_votes = Alsobrooks+Hogan+Scott) |>
  mutate(percent_alsobrooks = (Alsobrooks/total_votes)*100) |>
  mutate(percent_hogan = (Hogan/total_votes)*100) |>
  mutate(`Percent Difference` = percent_alsobrooks-percent_hogan)

#Draw the map
ggplot() +
  geom_sf(data = md_senate_county_24_with_map, aes(fill = `Percent Difference`)) +
  theme_minimal() +
  scale_fill_viridis_b(option="rocket") +
  labs(
    title="Alsobrooks Outrun Hogan in Prince George's and Most Counties in Central Maryland",
    subtitle = "According to the percentage difference in total votes between Alsobrooks and Hagan, Alsobrooks won 
more votes in central areas of Maryland, e.g. Prince George, Baltimore City and Montgomery.",
caption = "The scale represents the difference between Alsobrooks' and Hogan votes percentage of total votes.")
```

**Q2.** Make a county-level map showing the difference between Donald Trump's county-level performance this year and Larry Hogan's, using percentages to compare the two as you did in Q1. You can use the dataframe you initially loaded for Q1, and you'll need to load the 2024 presidential results in the data folder and join them to it before proceeding. Are there any counties where Trump got a higher percentage than Hogan? How would you describe the map showing the Trump-Hogan difference?

Also answer this: is a map the best way to present this data? What else could you make that might convey more information?

**A2.** Allegany was the county where Trump got a higher percentage than Hogan. I would describe the map as follows: Allegany is the only county in Maryland where Trump performed better than Hogan.

This map does a pretty good job of comparing the performance of Trump in difference counties, but if the main point of the story is to address how different Trump and Hogan are in each county, a stacked bars chart or group bars chart might be a more effective way to convey this information because they could show the difference between Trump and Hogan in a more direct and obvious way.

```{r}
#Download President Election Data
md_president_24 <- read_csv("data/md_pres_county_24.csv")

#Find Trump Percentage 
md_president_24 <- md_president_24 |>
  mutate(total_votes = Harris+Trump+Oliver+Stein+Kennedy) |>
  mutate(percent_trump = (Trump/total_votes)*100) |>
  select(County, percent_trump)

#Adjust Senate Data 
hogan_trump_with_map <- md_senate_county_24_with_map |>
  select(County, percent_hogan, geometry)

#Join the Dataframe find the difference
hogan_trump_with_map <- hogan_trump_with_map |>
  inner_join(md_president_24, join_by(County))
hogan_trump_with_map <- hogan_trump_with_map |>
  mutate(`Percent Difference` = percent_trump-percent_hogan)

#Draw the map
ggplot() +
  geom_sf(data = hogan_trump_with_map, aes(fill = `Percent Difference`)) +
  theme_minimal() +
  scale_fill_viridis_b(option="inferno") +
  labs(
    title="Trump Wins More than Hogan only in Allegany, Maryland",
    subtitle = "Trump gained more percentage votes in the presidential race than Hogan gained in the Senate race in 
only Allegany out of all counties in Maryland.",
caption = "The scale represents the difference between Trump's and Hogan's votes percentage of total votes.")

ggplot() +
  geom_dumbbell(data=hogan_trump_with_map, aes(y=County, x=percent_trump, xend=percent_hogan)) + theme_minimal()
```

**Q3** Make another map showing the difference between Larry Hogan's county-level performance this year and from his governor race in 2018, using percentages to compare the two as you did in Q2. You can use the dataframe you initially loaded for Q1, and you'll need to load the 2018 governor results in the data folder and join them to it before proceeding. Are there any counties where Hogan did better this year? How would you describe the map showing the difference?

**A3** There are no counties where Hogan gained more votes this year. I would describe the map that compared to his gubernatorial race in 2018, Hogan underperformed in every Maryland county for this year's Senate race.  

```{r}
#Download Hogan 2018 Election Data
md_gov_18 <- read_csv("data/md_gov_county_18.csv")

#Find Percentage of Hogan's Votes
md_gov_18 <- md_gov_18 |>
  mutate(total_votes = Hogan + Jealous + Quinn + Schlakman + `Write-ins`) |>
  mutate(percentage_hogan_18 = (Hogan/total_votes)*100) |>
  select(County, percentage_hogan_18)

#Adjust Senate 2024 Data 
hogan_18_and_24 <- md_senate_county_24_with_map |>
  rename(percent_hogan_24 = percent_hogan) |>
  select(County, percent_hogan_24, geometry)

#Join the Dataframe find the difference
hogan_18_and_24 <- hogan_18_and_24 |>
  inner_join(md_gov_18, join_by(County))
hogan_18_and_24 <- hogan_18_and_24 |>
  mutate(`Percent Difference` = percent_hogan_24-percentage_hogan_18)

#Draw the map
ggplot() +
  geom_sf(data = hogan_18_and_24, aes(fill = `Percent Difference`)) +
  theme_minimal() +
  scale_fill_viridis_b(option="viridis") +
  labs(
    title="Hogan Underperfroms This Year Compared to His Race in 2018",
    subtitle = "Hogan obtained votes for this year's U.S. Senate election fewer than those he got when running for 
2018 Governor race in every Maryland county.",
caption = "The scale represents the difference between 2020 and 2018 of Hogan's votes percentage of total votes.")
```

**Q4.** Choose your own map adventure! In the data folder I've included Maryland county-level results for the abortion rights amendment vote this year, plus the 2020 presidential results by county and 2022 governor's results by county. Using one of those and at least one other contest for comparison, make a county-level map showing what you think is the most interesting aspect of that data, and describe the results in terms of a possible story. You can bring in other data, such as Census information, if you like.

**A4.** I think the possible and interesting story here is about the 3 counties where Trump has done a better performance this year (Calvert, Washington, and Caroline) compared to his campaign in 2020. What are the factors that make these 3 areas different than other Counties, and what has been changed for the last 4 years?  

```{r}
#Download County Population and Map
md_county_map_and_population <- get_acs(geography = "county",
              variables = c(population = "B01001_001"),
              state = "MD",
              year = 2022,
              geometry = TRUE)
#Clean data 
md_county_map_and_population <- md_county_map_and_population |>
  rename(County = NAME) |>
  rename(population = estimate) |>
  mutate(County = str_to_title(str_remove_all(County,", Maryland|County"))) |>
  mutate(County = str_trim(County,side="both")) |>
  mutate(County = if_else(County == "St. Mary's", "Saint Mary's", County)) |> 
  mutate(County = if_else(County == "Baltimore", "Baltimore County", County)) |>
  select(County, population, geometry)

#Download County Population (I redo this because somehow I cannot get rid of the geometry column before I join the dataframe with President 2020 data, and it doesn't allow me to join 2 dataframe that have a geometry column.)
md_county_population <- get_acs(geography = "county",
              variables = c(population = "B01001_001"),
              state = "MD",
              year = 2022)
#Clean data 
md_county_population <- md_county_population |>
  rename(County = NAME) |>
  rename(population = estimate) |>
  mutate(County = str_to_title(str_remove_all(County,", Maryland|County"))) |>
  mutate(County = str_trim(County,side="both")) |>
  mutate(County = if_else(County == "St. Mary's", "Saint Mary's", County)) |> 
  mutate(County = if_else(County == "Baltimore", "Baltimore County", County)) |>
  select(County, population)

#Download President 2024 Data 
President_2024 <- read_csv("data/md_pres_county_24.csv")

#Join President 2024 Data with Map and Population + Find Trump Percent
Trump_2024 <- md_county_map_and_population |>
  inner_join(President_2024, join_by(County)) |>
  mutate(percent_trump_24 = (Trump/population)*100) |>
  select(County, percent_trump_24, geometry)

#Download President 2020 Data 
President_2020 <- read_csv("data/md_pres_county_20.csv")

#Join President 2020 Data with Map and Population + Find Trump Percent
Trump_2020 <- md_county_population |>
  inner_join(President_2020, join_by(County)) |>
  mutate(percent_trump_20 = (Trump/population)*100) |>
  select(County, percent_trump_20)

#Join Trump 2024 and 2020 to Find Percent Difference
Trump_2024_2020 <- Trump_2024 |>
  inner_join(Trump_2020, join_by(County)) |>
  mutate(`Percent Difference` = percent_trump_24-percent_trump_20)

#Draw the map
ggplot() +
  geom_sf(data = Trump_2024_2020, aes(fill = `Percent Difference`)) +
  theme_minimal() +
  scale_fill_viridis_b(option="cividis") +
  labs(
    title="Trump Beats His 2020 Race This Year in 3 Maryland Counties",
    subtitle = "Trump's 2024 presidential race outperformed his previous race in 2020 gaining more percent votes of
county's total population in 3 Maryland Counites including Calvert, Washington and Caroline.",
  caption = "The scale represents the difference between 2024 and 2020 in Trump votes percentage of the total population.")
```
